---
title: "Algorithm benchmarking script"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: hide
    theme: united
    highlight: tango
author: Anna Guadall
params:
  file: "/Users/anna/Documents/paper_cyto/scripts_and_files/files/beads_18_comp_50/ff.fcs"
  labels: "/Users/anna/Documents/paper_cyto/scripts_and_files/files/beads_18_comp_50/labels"
  outliers:
    label: "0: no outiliers, 1: outliers"
    value: 0
  diff: 
    labels: minimum difference among F1 scores that are not equivalent
    value: 0.05
  clusters: "5"
  dir: "dir"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, cache = T)
```

```{r libraries, message=FALSE}
library(FlowSOM)
library(caret)
library(cytofkit)
library(flowPeaks)
library(flowMeans)
library(ggplot2)
library(gridExtra)
library(umap)
library(knitr)
library(DepecheR)
```

# Reading the data
```{r data}
# import the fcs file
ff <- flowCore::read.FCS(params$file)

# Import the labels
labels <- readRDS(params$labels) # "labels" has c levels

# number of populations
n_pops <- length(levels(labels)) - params$outliers
```

```{r freqs}
# Computing  populations' frequencies
ref_pops_freq <- (prop.table(table(labels)))*100
table_frqs <- cbind(count(labels), round(ref_pops_freq, 2))
table_frqs <- table_frqs[,c(1:2,4)]
colnames(table_frqs) <- c("Labels", "Counts", "Frequencies")
```

Date: ``r Sys.Date()``  
File: ``r params$file``   
``r nrow(ff@exprs)`` events.  
`r n_pops` populations.  
`r ncol(ff@exprs)` parameters.  

```{r table freqs}
kable(table_frqs)
```

# Number of clusters (For `FlowSOM`)
```{r clusters}
clusters <- c(eval(parse(text = paste("c(", params$clusters, ")", sep = ""))))
```

# Functions
Normalize
```{r normalize}
normalize <- function(x) { return ((x - min(x)) / (max(x) - min(x))) }
```

Matching labels and predictions according the F1 score and merging populations that have been clustered together.
```{r matching}
matching <- function(clustering, labels){
  c <- table(clustering, labels) # Contingency matrix
  r <- prop.table(c, 2)          # Recall matrix
  p <- prop.table(c, 1)          # Precision matrix
  f <- 2 * p * r / (p + r)       # F1 matrix
  f[is.na(f)] <-  0              # replacing NAs by zeros
  
  # Finding the cell population (columns) 
  # with a higher score F1 for each cluster (rows):
  m <- apply(f, 1, function(x){which.max(x)})
  m <- as.list(m)
  
  # Normalize column maximums
  col_norm <- apply(f, 2, normalize)
  
  # F1 column equivalent maximums
  col_m <- apply(col_norm, 2,  function(x){ which(max(x)  - x < params$diff & x != 0) }  )
  col_m <- as.list(col_m)
  
  # Normalize row values (row-norm-F1)
  row_norm <- apply(f, 1, normalize) # why is the result transposed?
  row_norm <- t(row_norm)
  
  # Row-norm-F1 row equivalent maximums
  row_m <- apply( row_norm, 1, function(x){ which(max(x) - x < params$diff & x != 0) }  )
  row_m <- as.list(row_m)
  
  # MERGING POPULATIONS (IN CASE ONE CLUSTER MATCHES MANY POPULATIONS)
  merged_labels <- as.character(labels)
  
  count_merged_pops <- rep(1, length(labels))
  
  for (i in 1:length(labels)){
    for (j in unlist(col_m)){ 
      if(labels[i] %in% unlist(names(row_m[[j]]))){
        if(unlist(names(row_m[[j]]))[1] == names(col_m[unlist(names(row_m[[j]]))[1]])){
          merged_labels[i] <- paste( unlist( names(row_m[[j]]) ), collapse = "-" )
          count_merged_pops[i] <- length(unlist( names(row_m[[j]]) ))
        }
      }
    }
  }
  
  
  # MATCHING CLUSTERS TO LABELS
  # Empty list 
  matched_merged_clusters <- rep("NA", length(clustering))
  matched_clusters <- rep("NA", length(clustering))
  
  # Replacing the numbers of the clusters by the names of the cell types:
  for(i in 1:length(clustering)){
    for(j in 1:length(row_m)){ 
      if(clustering[i] == names(row_m)[j]){ # if  a cluster number equals a row maximum...
        # ... give it the name of the absolute row maximum
        matched_clusters[i] <- levels(labels)[m[[j]]]
        # ... give it the name of the corresponding populations,
        # merging the name to the other equivalent row maximums, if there are
        matched_merged_clusters[i] <- paste(levels(labels)[as.numeric(unlist(row_m[[j]]))], 
                                            collapse = "-")
      }
    }
  }
  
  # Number of PARTTIONS (matched clusters) 
  partitions <- length(table(matched_clusters))
  
  # Factorize matched (merged) clusters and labels 
  merged_labels <- as.factor(merged_labels)
  matched_merged_clusters <- factor(matched_merged_clusters, levels = levels(merged_labels)) 
  matched_clusters <- factor(matched_clusters, levels = levels(labels)) 
  # same level order than merged_labels
  # adds labels that have not been predicted
  
  matched <- list("c" = c, "f" = f, "m" = m, "merged_labels" = merged_labels,
                  "clusters" = matched_clusters, "merged_clusters" = matched_merged_clusters,
                  "partitions" = partitions, "count_merged_pops" = count_merged_pops)
  return(matched)
}
```

Computing DIFFERENT MEAN F1 scores:
```{r f1}
mean_f1 <- function(cm, cm_merged, merged_labels, count_merged_pops){
  # NOT MERGED, ALL
  f1_list <- cm$byClass[,"F1"]
  f1_zeros <- f1_list
  f1_zeros <- ifelse(is.na(f1_list), 0, f1_list)
  mf1 <- mean(f1_zeros) # MEAN
  
  # NOT MERGED, ALL, WEIGHTED
  pop_freq <- prop.table(table(labels))
  mf1_w <- t(pop_freq) %*% f1_zeros 
  
  # NOT MERGED, ALL, INVERSED WEIGHTS
  inv_freq <- prop.table(1/table(labels))
  mf1_i_w <- t(inv_freq) %*% f1_zeros 
  
  # MERGED, ALL
  merged_f1_list <- cm_merged$byClass[,"F1"]
  merged_f1_zeros <- merged_f1_list
  merged_f1_zeros <- ifelse(is.na(merged_f1_list), 0, merged_f1_list)
  mf1_m <- mean(merged_f1_zeros) # MEAN
  
  # MERGED, CORRECTED
  merged_table <- table(merged_labels, count_merged_pops)
  merged_table <- cbind(merged_table, correction = 0)
  
  for(i in 1:nrow(merged_table)){
    for(j in 1:(ncol(merged_table)-1)){
      if(merged_table[i, j] != 0){
        merged_table[i, "correction"] <- merged_f1_zeros[i]/as.numeric(colnames(merged_table)[j])
      }
    }
  }
  
  mf1_m_all_corrected <- mean(merged_table[,"correction"]) # CORRECTED MEAN
  
  f1 <- list("mf1" = mf1, "mf1_w" = mf1_w, "mf1_i_w" = mf1_i_w,
             "mf1_m" = mf1_m, "mf1_m_c" = mf1_m_all_corrected)
  
  return(f1)
}
```

# Preparing results tables
```{r table}
pops_f1 <- c()
summ <- data.frame(File = NA, n = 0, Method = NA, Clusters = 0, Partitions = 0, 
                   Mean_F1 = 0, Weighted_mean_F1 = 0, Inversed_weights_mean_F1 = 0,
                   Mean_F1_merged = 0, Corrected_mean_F1_merged = 0,
                   Reduction_user_time = 0, Reduction_elapsed_time = 0,
                   Clustering_user_time = 0, Clustering_elapsed_time = 0)
```

# t-SNE dimensionality reduction
```{r tsne}
ptm <- proc.time() # measuring CPU time
tsne_red <-  cytof_dimReduction(data = ff@exprs, method = "tsne", tsneSeed = 42)
tsne_time <- proc.time() - ptm

tsne_red <- as.data.frame(tsne_red)
```

```{r tsne plot, fig.width = 12}
tsne_df <- cbind(as.data.frame(tsne_red), as.data.frame(labels))
coloring <-  c("darkgoldenrod1", "blue4", "blueviolet", "red", "turquoise", 
               "burlywood3", "cadetblue", "chartreuse", "chartreuse4", "chocolate1",
               "plum1", "cornflowerblue", "yellow", "cyan", "blue",
               "darkgrey","khaki1", "deeppink1", "deeppink4", "black")
cytof_clusterPlot(data = tsne_df, xlab = "tsne_1", ylab = "tsne_2", point_size = 0.5, labelRepel = F,
                  cluster = "labels", sampleLabel = FALSE, labelSize = 0, title = "Labels",
                  clusterColor = coloring[1:n_pops])
```

```{r tsne plot png}
png("tsne.png", res=150, width = 2000, height = 1000)
cytof_clusterPlot(data = tsne_df, xlab = "tsne_1", ylab = "tsne_2", point_size = 0.5, labelRepel = F,
                  cluster = "labels", sampleLabel = FALSE, labelSize = 0, title = "Labels",
                  clusterColor = coloring[1:n_pops])
dev.off()
```

# UMAP dimensionality reduction
```{r umap}
ptm <- proc.time() # measuring CPU time
umap_red <- umap(ff@exprs, random_state = 42)
umap_time <- proc.time() - ptm

colnames(umap_red$layout) <- c("UMAP_1", "UMAP_2")
```

```{r umap plot, fig.width = 12}
# preparing data for plotting on UMAP:
umap_df <- cbind(as.data.frame(umap_red$layout), as.data.frame(labels))
cytof_clusterPlot(data = umap_df, xlab = "UMAP_1", ylab = "UMAP_2", point_size = 0.5, labelRepel = F,
                  cluster = "labels", sampleLabel = FALSE, labelSize = 0, title = "Labels",
                  clusterColor = coloring[1:n_pops])
```

```{r umap plot png}
png("umap.png", res=150, width = 2000, height = 1000)
cytof_clusterPlot(data = umap_df, xlab = "UMAP_1", ylab = "UMAP_2", point_size = 0.5, labelRepel = F,
                  cluster = "labels", sampleLabel = FALSE, labelSize = 0, title = "Labels",
                  clusterColor = coloring[1:n_pops])
dev.off()
```

# Clustering on all dimensions
## `FlowSOM`
```{r flowsom}
# Parameters
method <- "FlowSOM"

for(i in 1:length(clusters)){
  # META-CLUSTERING
  set.seed(42)
  
  ptm <- proc.time() # measuring CPU time
  fs <- FlowSOM(ff, compensate = F,transform = F, scale = F,
                colsToUse = colnames(ff@exprs), nClus = clusters[i], seed = 42)
  ptm <- proc.time() - ptm
  
  clustering <- fs$metaclustering[fs$FlowSOM$map$mapping[,1]]
  
  # MATCHING LABELS AND PREDICTIONS
  matched <- matching(clustering, labels)
  
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                        reference = labels[labels != "outliers"])
  cm_merged <- confusionMatrix(
    data = matched$merged_clusters[matched$merged_labels != "outliers"],
    reference = matched$merged_labels[matched$merged_labels != "outliers"]
    )
  
  # COMPUTE MEAN F1 and WEIGHTED MEAN F1
  mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
  # Storing the result 
  summ[i,c(1,3)] <- c(params$file, method)
  summ[i,-c(1,3)] <- c(nrow(ff@exprs), clusters[i], matched$partitions, 
                       mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                       NA, NA, ptm[1] + ptm[4], ptm[3])
  
  # FOR VISUALIZATION
  cat("\n", method, clusters[i], "clusters\n")
  cat("\nContingency matrix")
  print(matched$c)
  cat("\nF1 matrix")
  print(round(matched$f, 2))
  cat("\nConfusion matrix")
  print(cm$table)
  cat("\nConfusion matrix (MERGED)")
  print(cm_merged$table)
}
```

## RphenoGraph
```{r phenograph}
method <- "PhenoGraph"

# clustering
ptm <- proc.time() # measuring CPU time
pred <- Rphenograph(ff@exprs) # default k = 30 nearest neighbours
ptm <- proc.time() - ptm

clustering <- pred$membership

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = labels[labels != "outliers"])
cm_merged <- confusionMatrix(
  data = matched$merged_clusters[matched$merged_labels != "outliers"],
  reference = matched$merged_labels[matched$merged_labels != "outliers"]
)

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
summ[nrow(summ)+1,c(1,3)] <- c(params$file, method)
summ[nrow(summ),-c(1,3)] <- c(nrow(ff@exprs), length(table(clustering)),
                              matched$partitions,
                              mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                              NA, NA, ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
cat("\n", method)
cat("\nContingency matrix")
print(matched$c)
cat("\nF1 matrix")
print(round(matched$f, 2))
cat("\nConfusion matrix")
print(cm$table)
cat("\nConfusion matrix (MERGED)")
print(cm_merged$table)
```

## flowMeans
```{r flowmeans}
method <- "flowMeans"

# clustering
ptm <- proc.time() # measuring CPU time
pred <- flowMeans(ff@exprs, MaxN = n_pops )
ptm <- proc.time() - ptm

clustering <- pred@Labels[[1]]

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = labels[labels != "outliers"])
cm_merged <- confusionMatrix(
  data = matched$merged_clusters[matched$merged_labels != "outliers"],
  reference = matched$merged_labels[matched$merged_labels != "outliers"]
)

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
summ[nrow(summ)+1,c(1,3)] <- c(params$file, method)
summ[nrow(summ),-c(1,3)] <- c(nrow(ff@exprs), length(table(clustering)),
                              matched$partitions,
                              mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                              NA, NA, ptm[1] + ptm[4], ptm[3])
# FOR VISUALIZATION
cat("\n", method)
cat("\nContingency matrix")
print(matched$c)
cat("\nF1 matrix")
print(round(matched$f, 2))
cat("\nConfusion matrix")
print(cm$table)
cat("\nConfusion matrix (MERGED)")
print(cm_merged$table)
```

## flowPeaks
```{r flowpeaks}
method <- "flowPeaks"

# clustering
ptm <- proc.time() # measuring CPU time
pred <- flowPeaks(ff@exprs)
ptm <- proc.time() - ptm

clustering <- pred$peaks.cluster

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = labels[labels != "outliers"])
cm_merged <- confusionMatrix(
  data = matched$merged_clusters[matched$merged_labels != "outliers"],
  reference = matched$merged_labels[matched$merged_labels != "outliers"]
)

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
summ[nrow(summ)+1, c(1,3)] <- c(params$file, method)
summ[nrow(summ), -c(1,3)] <- c(nrow(ff@exprs), length(table(clustering)),
                              matched$partitions,
                              mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                              NA, NA, ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
cat("\n", method)
cat("\nContingency matrix")
print(matched$c)
cat("\nF1 matrix")
print(round(matched$f, 2))
cat("\nConfusion matrix")
print(cm$table)
cat("\nConfusion matrix (MERGED)")
print(cm_merged$table)
```

## DEPECHE
```{r depeche}
method <- "DEPECHE"

# clustering
ptm <- proc.time() # measuring CPU time
pred <-  depeche(ff@exprs, createOutput = F)
ptm <- proc.time() - ptm

clustering <- pred$clusterVector

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = labels[labels != "outliers"])
cm_merged <- confusionMatrix(
  data = matched$merged_clusters[matched$merged_labels != "outliers"],
  reference = matched$merged_labels[matched$merged_labels != "outliers"]
)

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
summ[nrow(summ)+1, c(1,3)] <- c(params$file, method)
summ[nrow(summ), -c(1,3)] <- c(nrow(ff@exprs), length(table(clustering)),
                              matched$partitions,
                              mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                              NA, NA, ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
cat("\n", method)
cat("\nContingency matrix")
print(matched$c)
cat("\nF1 matrix")
print(round(matched$f, 2))
cat("\nConfusion matrix")
print(cm$table)
cat("\nConfusion matrix (MERGED)")
print(cm_merged$table)
```

# Clustering on t-SNE-reduced data (2D)
## RphenoGraph
```{r tsne phenograph}
method <- "t-SNE + PhenoGraph"

# clustering
ptm <- proc.time() # measuring CPU time
pred <- Rphenograph(tsne_red) # default k = 30 nearest neighbours
ptm <- proc.time() - ptm

clustering <- pred$membership

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = labels[labels != "outliers"])
cm_merged <- confusionMatrix(
  data = matched$merged_clusters[matched$merged_labels != "outliers"],
  reference = matched$merged_labels[matched$merged_labels != "outliers"]
)

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
summ[nrow(summ)+1, c(1,3)] <- c(params$file, method)
summ[nrow(summ), -c(1,3)] <- c(nrow(ff@exprs), length(table(clustering)),
                               matched$partitions,
                               mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                               tsne_time[1] + tsne_time[4], tsne_time[3], ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
cat("\n", method)
cat("\nContingency matrix")
print(matched$c)
cat("\nF1 matrix")
print(round(matched$f, 2))
cat("\nConfusion matrix")
print(cm$table)
cat("\nConfusion matrix (MERGED)")
print(cm_merged$table)
```

## flowMeans
```{r tsne flowmeans}
method <- "t-SNE + flowMeans"

# clustering
ptm <- proc.time() # measuring CPU time
pred <- flowMeans(tsne_red, MaxN = n_pops )
ptm <- proc.time() - ptm

clustering <- pred@Labels[[1]]

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = labels[labels != "outliers"])
cm_merged <- confusionMatrix(
  data = matched$merged_clusters[matched$merged_labels != "outliers"],
  reference = matched$merged_labels[matched$merged_labels != "outliers"]
)

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
summ[nrow(summ)+1, c(1,3)] <- c(params$file, method)
summ[nrow(summ), -c(1,3)] <- c(nrow(ff@exprs), length(table(clustering)),
                               matched$partitions,
                               mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                               tsne_time[1] + tsne_time[4], tsne_time[3], ptm[1] + ptm[4], ptm[3])


# FOR VISUALIZATION
cat("\n", method)
cat("\nContingency matrix")
print(matched$c)
cat("\nF1 matrix")
print(round(matched$f, 2))
cat("\nConfusion matrix")
print(cm$table)
cat("\nConfusion matrix (MERGED)")
print(cm_merged$table)
```

## flowPeaks
```{r tsne flowpeaks}
method <- "t-SNE + flowPeaks"

# clustering
ptm <- proc.time() # measuring CPU time
pred <- flowPeaks(tsne_red) 
ptm <- proc.time() - ptm

clustering <- pred$peaks.cluster

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = labels[labels != "outliers"])
cm_merged <- confusionMatrix(
  data = matched$merged_clusters[matched$merged_labels != "outliers"],
  reference = matched$merged_labels[matched$merged_labels != "outliers"]
)

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
summ[nrow(summ)+1, c(1,3)] <- c(params$file, method)
summ[nrow(summ), -c(1,3)] <- c(nrow(ff@exprs), length(table(clustering)),
                               matched$partitions,
                               mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                               tsne_time[1] + tsne_time[4], tsne_time[3], ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
cat("\n", method)
cat("\nContingency matrix")
print(matched$c)
cat("\nF1 matrix")
print(round(matched$f, 2))
cat("\nConfusion matrix")
print(cm$table)
cat("\nConfusion matrix (MERGED)")
print(cm_merged$table)
```

## DEPECHE
```{r tsne depeche}
method <- "t-SNE + DEPECHE"

# clustering
ptm <- proc.time() # measuring CPU time
pred <-  depeche(tsne_red, createOutput = F)
ptm <- proc.time() - ptm

clustering <- pred$clusterVector

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = labels[labels != "outliers"])
cm_merged <- confusionMatrix(
  data = matched$merged_clusters[matched$merged_labels != "outliers"],
  reference = matched$merged_labels[matched$merged_labels != "outliers"]
)

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
summ[nrow(summ)+1, c(1,3)] <- c(params$file, method)
summ[nrow(summ), -c(1,3)] <- c(nrow(ff@exprs), length(table(clustering)),
                               matched$partitions,
                               mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                               tsne_time[1] + tsne_time[4], tsne_time[3], ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
cat("\n", method)
cat("\nContingency matrix")
print(matched$c)
cat("\nF1 matrix")
print(round(matched$f, 2))
cat("\nConfusion matrix")
print(cm$table)
cat("\nConfusion matrix (MERGED)")
print(cm_merged$table)
```

## ClusterX
```{r tsne clusterx}
method <- "t-SNE + ClusterX"

# clustering
ptm <- proc.time() # measuring CPU time
clustering <- cytof_cluster(tsne_red, method = "ClusterX")
ptm <- proc.time() - ptm

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = labels[labels != "outliers"])
cm_merged <- confusionMatrix(
  data = matched$merged_clusters[matched$merged_labels != "outliers"],
  reference = matched$merged_labels[matched$merged_labels != "outliers"]
)

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
summ[nrow(summ)+1,] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                         matched$partitions,
                         mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                         tsne_time[1] + tsne_time[4], tsne_time[3], ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
cat("\n", method)
cat("\nContingency matrix")
print(matched$c)
cat("\nF1 matrix")
print(round(matched$f, 2))
cat("\nConfusion matrix")
print(cm$table)
cat("\nConfusion matrix (MERGED)")
print(cm_merged$table)
```

# Clustering on UMAP-reduced data (2D)
## RphenoGraph
```{r umap phenograph}
method <- "UMAP + PhenoGraph"

# clustering
ptm <- proc.time() # measuring CPU time
pred <- Rphenograph(umap_red$layout) # default k = 30 nearest neighbours
ptm <- proc.time() - ptm

clustering <- pred$membership

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = labels[labels != "outliers"])
cm_merged <- confusionMatrix(
  data = matched$merged_clusters[matched$merged_labels != "outliers"],
  reference = matched$merged_labels[matched$merged_labels != "outliers"]
)

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
summ[nrow(summ)+1,] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                         matched$partitions,
                         mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                         umap_time[1] + umap_time[4], umap_time[3], ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
cat("\n", method)
cat("\nContingency matrix")
print(matched$c)
cat("\nF1 matrix")
print(round(matched$f, 2))
cat("\nConfusion matrix")
print(cm$table)
cat("\nConfusion matrix (MERGED)")
print(cm_merged$table)
```

## flowMeans
```{r umap flowmeans}
method <- "UMAP + flowMeans"

# clustering
ptm <- proc.time() # measuring CPU time
pred <- flowMeans(umap_red$layout, MaxN = n_pops )
ptm <- proc.time() - ptm

clustering <- pred@Labels[[1]]

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = labels[labels != "outliers"])
cm_merged <- confusionMatrix(
  data = matched$merged_clusters[matched$merged_labels != "outliers"],
  reference = matched$merged_labels[matched$merged_labels != "outliers"]
)

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
summ[nrow(summ)+1,] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                         matched$partitions,
                         mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                         umap_time[1] + umap_time[4], umap_time[3], ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
cat("\n", method)
cat("\nContingency matrix")
print(matched$c)
cat("\nF1 matrix")
print(round(matched$f, 2))
cat("\nConfusion matrix")
print(cm$table)
cat("\nConfusion matrix (MERGED)")
print(cm_merged$table)
```

## flowPeaks
```{r umap flowpeaks}
method <- "UMAP + flowPeaks"

# clustering
ptm <- proc.time() # measuring CPU time
pred <- flowPeaks(umap_red$layout)
ptm <- proc.time() - ptm

clustering <- pred$peaks.cluster

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = labels[labels != "outliers"])
cm_merged <- confusionMatrix(
  data = matched$merged_clusters[matched$merged_labels != "outliers"],
  reference = matched$merged_labels[matched$merged_labels != "outliers"]
)

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
summ[nrow(summ)+1,] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                         matched$partitions,
                         mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                         umap_time[1] + umap_time[4], umap_time[3], ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
cat("\n", method)
cat("\nContingency matrix")
print(matched$c)
cat("\nF1 matrix")
print(round(matched$f, 2))
cat("\nConfusion matrix")
print(cm$table)
cat("\nConfusion matrix (MERGED)")
print(cm_merged$table)
```

## DEPECHE
```{r umap depeche}
method <- "UMAP + DEPECHE"

# clustering
ptm <- proc.time() # measuring CPU time
pred <-  depeche(umap_red$layout, createOutput = F)
ptm <- proc.time() - ptm

clustering <- pred$clusterVector

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = labels[labels != "outliers"])
cm_merged <- confusionMatrix(
  data = matched$merged_clusters[matched$merged_labels != "outliers"],
  reference = matched$merged_labels[matched$merged_labels != "outliers"]
)

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
summ[nrow(summ)+1,] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                         matched$partitions,
                         mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                         umap_time[1] + umap_time[4], umap_time[3], ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
cat("\n", method)
cat("\nContingency matrix")
print(matched$c)
cat("\nF1 matrix")
print(round(matched$f, 2))
cat("\nConfusion matrix")
print(cm$table)
cat("\nConfusion matrix (MERGED)")
print(cm_merged$table)
```

## ClusterX
```{r umap clusterx}
method <- "UMAP + ClusterX"

# clustering
ptm <- proc.time() # measuring CPU time
clustering <- cytof_cluster(umap_red$layout, method = "ClusterX")
ptm <- proc.time() - ptm

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = labels[labels != "outliers"])
cm_merged <- confusionMatrix(
  data = matched$merged_clusters[matched$merged_labels != "outliers"],
  reference = matched$merged_labels[matched$merged_labels != "outliers"]
)

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
summ[nrow(summ)+1,] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                         matched$partitions,
                         mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                         umap_time[1] + umap_time[4], umap_time[3], ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
cat("\n", method)
cat("\nContingency matrix")
print(matched$c)
cat("\nF1 matrix")
print(round(matched$f, 2))
cat("\nConfusion matrix")
print(cm$table)
cat("\nConfusion matrix (MERGED)")
print(cm_merged$table)
```

# Results
```{r results}
library(knitr)
for(i in 7:15){
  summ[,i] <- as.numeric(summ[,i])
}

kable(summ, digits = 3)
```

```{r}
# export results (mean F1)
write.table(summ, 
            file =  file.path("files", params$dir, "summary.txt"), append = T, sep = "\t", row.names = F, col.names = T)
```

# Libraries
```{r}
sessionInfo()
```













