---
title: "Algorithm benchmarking script"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: hide
    theme: united
    highlight: tango
author: Anna Guadall
params:
  file: "ff_03_2019-04-16.fcs"
  labels: "labels_03_2019-04-16"
  outliers:
    label: "0: no outiliers, 1: outliers"
    value: 0
  diff : 
    labels: minimum difference among F1 scores that are not equivalent
    value: 0.5
  clusters: "5,15,20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, cache = T)
```

```{r libraries, message=FALSE}
library(FlowSOM)
library(caret)
library(cytofkit)
library(flowPeaks)
library(flowMeans)
library(ggplot2)
library(gridExtra)
library(umap)
library(knitr)
library(DepecheR)
```

# Reading the data
```{r data}
# import the fcs file
ff <- flowCore::read.FCS(params$file)

# Import the labels
labels <- readRDS(params$labels) # "labels" has c levels

# number of populations
n_pops <- length(levels(labels)) - params$outliers
```

```{r freqs}
# Computing  populations' frequencies
(ref_pops_freq <- (prop.table(table(labels)))*100)
cbind(count(labels), round(ref_pops_freq, 2))
```

File: ``r params$file``   
``r nrow(ff@exprs)`` cells.  
`r n_pops` populations.  
`r ncol(ff@exprs)` parameters.  

# Number of clusters (For `FlowSOM`)
```{r clusters}
clusters <- c(eval(parse(text = paste("c(", params$clusters, ")", sep = ""))))
```

# Functions
Matching labels and predictions according the F1 score and merging populations that have been clustered together.
```{r matching}
matching <- function(clustering, labels){
  c <- table(clustering, labels) # Contingency matrix
  r <- prop.table(c, 2)          # Recall matrix
  p <- prop.table(c, 1)          # Precision matrix
  f <- 2 * p * r / (p + r)       # F1 matrix
  f[is.na(f)] <-  0              # replacing NAs by zeroes
  
  # Finding the cell population (columns) 
  # with a higher number of cells for each cluster (rows):
  m <- apply(f, 1, function(x){which.max(x)})
  m <- as.list(m)
  
  # Column equivalent maximums
  col_m <- apply(f, 2,  function(x){ which(max(x)  - x < params$diff & x != 0) }  )
  col_m <- as.list(col_m)
  
  # Row equivalent maximums
  row_m <- apply( f, 1, function(x){ which(max(x) - x < params$diff & x != 0) }  )
  row_m <- as.list(row_m)
  
  # MERGING POPULATIONS (IN CASE ONE CLUSTER MATCHES MANY POPULATIONS)
  matched_labels <- as.character(labels)
  
  for (i in 1:length(labels)){
    for (j in unlist(col_m)){ 
      if(labels[i] %in% unlist(names(row_m[[j]]))){
        if(unlist(names(row_m[[j]]))[1] == names(col_m[unlist(names(row_m[[j]]))[1]])){
          matched_labels[i] <- paste( unlist( names(row_m[[j]]) ), collapse = "-" )
        }
      }
    }
  }
  
  
  # MATCHING CLUSTERS TO LABELS
  # Empty list 
  matched_clusters <- rep("NA", length(clustering))
  
  # Replacing the numbers of the clusters by the names of the cell types:
  for(i in 1:length(clustering)){
    for(j in 1:length(row_m)){ 
      if(clustering[i] == names(row_m)[j]){ # if cluster numbers equals a row maximum...
        if(clustering[i] %in% unlist(col_m)){ # # ... that it is also a column maximum...
          # ... give it the name of the corresponding populations,
          # merging the name to the other equivalent row maximums, if it is the case
          matched_clusters[i] <- paste(levels(labels)[as.numeric(unlist(row_m[[j]]))], 
                                       collapse = "-")
          
        } else { # if the cluster number equals a row maximum that is not a column maximum...
          #... give it the name of the corresponding population
          matched_clusters[i] <- levels(labels)[m[[j]]]
        }
      }
    }
  }
  

  # Factorize matched labnels and clusters 
  matched_labels <- as.factor(matched_labels)
  # Give clusters the same level order than matched_labels
  matched_clusters <- factor(matched_clusters, levels = levels(matched_labels)) 
  
  # MATCHED POPULATIONS FREQUENCIES
  freq <- prop.table(table(matched_labels))
  
  matched <- list("c" = c, "m" = m, "clusters" = matched_clusters, "labels" = matched_labels,
                  "freqs" = freq)
  return(matched)
}
```

Computing DIFFERENT MEAN F1 scores:
```{r f1}
mean_f1 <- function(cm, pop_freq){
  f1_list <- cm$byClass[,"F1"] # cm: confusion matrix
  
  f1_no_nas <- f1_list[!is.na(f1_list)] # removing NAs from F1 list
  pop_freq_no_nas <- pop_freq[!is.na(f1_list)] / sum(pop_freq[!is.na(f1_list)])
  mean_f1_found <- mean(f1_no_nas) # MEAN F1 (FOUND POPS) 
  w_mean_f1_found <- t(pop_freq_no_nas) %*% f1_no_nas # WEIGHTED MEAN F1 (FOUND POPS)
  
  f1_list[is.na(f1_list)] <-  0 # replacing NAs by zero
  mean_f1_all <- mean(f1_list)   # MEAN F1 (ALL POPS)
  w_mean_f1_all <- t(pop_freq) %*% f1_list # WEIGHTED MEAN F1 (ALL POPS)
  
  f1 <- list("found" = mean_f1_found, "w_found" = w_mean_f1_found, 
             "all" = mean_f1_all, "w_all" = w_mean_f1_all)
  return(f1)
}
```

# Preparing results tables
```{r table}
pops_f1 <- c()
summ <- data.frame(File = NA, n = 0, Method = NA, Clusters = 0, Partitions = 0, 
                   Mean_F1_found = 0, Weighted_mean_F1_found = 0, Mean_F1_all = 0, Weighted_mean_F1_all = 0,
                   Reduction_user_time = 0, Reduction_elapsed_time = 0,
                   Clustering_user_time = 0, Clustering_elapsed_time = 0)
```

# t-SNE dimensionality reduction
```{r tsne}
ptm <- proc.time() # measuring CPU time
tsne_red <-  cytof_dimReduction(data = ff@exprs, method = "tsne", tsneSeed = 42)
tsne_time <- proc.time() - ptm
```

# UMAP dimensionality reduction
```{r umap}
ptm <- proc.time() # measuring CPU time
umap_red <- umap(ff@exprs, random_state = 42)
umap_time <- proc.time() - ptm

colnames(umap_red$layout) <- c("UMAP_1", "UMAP_2")
```

# Clustering on all dimensions
## `FlowSOM`
```{r flowsom}
# Parameters
method <- "FlowSOM"

for(i in 1:length(clusters)){
  # META-CLUSTERING
  set.seed(42)
  
  ptm <- proc.time() # measuring CPU time
  fs <- FlowSOM(ff, compensate = F,transform = F, scale = F,
                colsToUse = colnames(ff@exprs), nClus = clusters[i], seed = 42)
  ptm <- proc.time() - ptm
  
  clustering <- fs$metaclustering[fs$FlowSOM$map$mapping[,1]]
  
  # MATCHING LABELS AND PREDICTIONS
  matched <- matching(clustering, labels)
  number_partitions <- length(table(matched$clusters))
  
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                        reference = matched$labels[labels != "outliers"])
  # Add the F1 values to a list 
  pops_f1 <- c(pops_f1, cm$byClass[, "F1"])
  
  # COMPUTE MEAN F1 and WEIGHTED MEAN F1
  mf1 <- mean_f1(cm, matched$freqs)
  # Storing the result 
  summ[i,] <- c(params$file, nrow(ff@exprs), method, clusters[i], number_partitions, 
                mf1$found, mf1$w_found, mf1$all, mf1$w_all, NA, NA, ptm[1] + ptm[4], ptm[3])

  # FOR VISUALIZATION
  # Print the unmatched contingency matrix:
  cat("\n", method, clusters[i], "clusters\n")
  print(matched$c)
  # Print the matched confusion matrix:
  print(cm$table)
}
```

## RphenoGraph
```{r phenograph}
method <- "PhenoGraph"

# clustering
ptm <- proc.time() # measuring CPU time
pred <- Rphenograph(ff@exprs) # default k = 30 nearest neighbours
ptm <- proc.time() - ptm

clustering <- pred$membership

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)
number_partitions <- length(table(matched$clusters))

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = matched$labels[labels != "outliers"])
# Add the F1 values to a list 
pops_f1 <- c(pops_f1, cm$byClass[, "F1"])

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, matched$freqs)
# Storing the result 
summ[(nrow(summ)+1),] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                           number_partitions, mf1$found, mf1$w_found, mf1$all, mf1$w_all,
                           NA, NA, ptm[1] + ptm[4], ptm[3])
# FOR VISUALIZATION
# Print the unmatched contingency matrix:
cat("\n", method, "\n")
print(matched$c)
# Print the matched confusion matrix:
print(cm$table)
```

## flowMeans
```{r flowmeans}
method <- "flowMeans"

# clustering
ptm <- proc.time() # measuring CPU time
pred <- flowMeans(ff@exprs, MaxN = n_pops )
ptm <- proc.time() - ptm

clustering <- pred@Labels[[1]]

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)
number_partitions <- length(table(matched$clusters))

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = matched$labels[labels != "outliers"])
# Add the F1 values to a list 
pops_f1 <- c(pops_f1, cm$byClass[, "F1"])

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, matched$freqs)
# Storing the result 
summ[(nrow(summ)+1),] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                           number_partitions, mf1$found, mf1$w_found, mf1$all, mf1$w_all,
                           NA, NA, ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
# Print the unmatched contingency matrix:
cat("\n", method, "\n")
print(matched$c)
# Print the matched confusion matrix:
print(cm$table)
```

## flowPeaks
```{r flowpeaks}
method <- "flowPeaks"

# clustering
ptm <- proc.time() # measuring CPU time
pred <- flowPeaks(ff@exprs)
ptm <- proc.time() - ptm

clustering <- pred$peaks.cluster

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)
number_partitions <- length(table(matched$clusters))

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = matched$labels[labels != "outliers"])
# Add the F1 values to a list 
pops_f1 <- c(pops_f1, cm$byClass[, "F1"])

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, matched$freqs)
# Storing the result 
summ[(nrow(summ)+1),] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                           number_partitions, mf1$found, mf1$w_found, mf1$all, mf1$w_all,
                           NA, NA, ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
# Print the unmatched contingency matrix:
cat("\n", method, "\n")
print(matched$c)
# Print the matched confusion matrix:
print(cm$table)
```

## DEPECHE
```{r depeche}
method <- "DEPECHE"

# clustering
ptm <- proc.time() # measuring CPU time
pred <-  depeche(ff@exprs)
ptm <- proc.time() - ptm

clustering <- pred$clusterVector

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)
number_partitions <- length(table(matched$clusters))

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = matched$labels[labels != "outliers"])
# Add the F1 values to a list 
pops_f1 <- c(pops_f1, cm$byClass[, "F1"])

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, matched$freqs)
# Storing the result 
summ[(nrow(summ)+1),] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                           number_partitions, mf1$found, mf1$w_found, mf1$all, mf1$w_all,
                           NA, NA, ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
# Print the unmatched contingency matrix:
cat("\n", method, "\n")
print(matched$c)
# Print the matched confusion matrix:
print(cm$table)
```

# Clustering on t-SNE-reduced data (2D)
## RphenoGraph
```{r tsne phenograph}
method <- "t-SNE + PhenoGraph"

# clustering
ptm <- proc.time() # measuring CPU time
pred <- Rphenograph(tsne_red) # default k = 30 nearest neighbours
ptm <- proc.time() - ptm

clustering <- pred$membership

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)
number_partitions <- length(table(matched$clusters))

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = matched$labels[labels != "outliers"])
# Add the F1 values to a list 
pops_f1 <- c(pops_f1, cm$byClass[, "F1"])

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, matched$freqs)
# Storing the result 
summ[(nrow(summ)+1),] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                           number_partitions, mf1$found, mf1$w_found, mf1$all, mf1$w_all,
                           tsne_time[1] + tsne_time[4], tsne_time[3], ptm[1] + ptm[4], ptm[3])
# FOR VISUALIZATION
# Print the unmatched contingency matrix:
cat("\n", method, "\n")
print(matched$c)
# Print the matched confusion matrix:
print(cm$table)
```

## flowMeans
```{r tsne flowmeans}
method <- "t-SNE + flowMeans"

# clustering
ptm <- proc.time() # measuring CPU time
pred <- flowMeans(tsne_red, MaxN = n_pops )
ptm <- proc.time() - ptm

clustering <- pred@Labels[[1]]

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)
number_partitions <- length(table(matched$clusters))

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = matched$labels[labels != "outliers"])
# Add the F1 values to a list 
pops_f1 <- c(pops_f1, cm$byClass[, "F1"])

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, matched$freqs)
# Storing the result 
summ[(nrow(summ)+1),] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                           number_partitions, mf1$found, mf1$w_found, mf1$all, mf1$w_all,
                           tsne_time[1] + tsne_time[4], tsne_time[3], ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
# Print the unmatched contingency matrix:
cat("\n", method, "\n")
print(matched$c)
# Print the matched confusion matrix:
print(cm$table)
```

## flowPeaks
```{r tsne flowpeaks}
method <- "t-SNE + flowPeaks"

# clustering
ptm <- proc.time() # measuring CPU time
pred <- flowPeaks(tsne_red) # default k = 30 nearest neighbours
ptm <- proc.time() - ptm

clustering <- pred$peaks.cluster

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)
number_partitions <- length(table(matched$clusters))

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = matched$labels[labels != "outliers"])
# Add the F1 values to a list 
pops_f1 <- c(pops_f1, cm$byClass[, "F1"])

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, matched$freqs)
# Storing the result 
summ[(nrow(summ)+1),] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                           number_partitions, mf1$found, mf1$w_found, mf1$all, mf1$w_all,
                           tsne_time[1] + tsne_time[4], tsne_time[3], ptm[1] + ptm[4], ptm[3])
# FOR VISUALIZATION
# Print the unmatched contingency matrix:
cat("\n", method, "\n")
print(matched$c)
# Print the matched confusion matrix:
print(cm$table)
```

## DEPECHE
```{r tsne depeche}
method <- "t-SNE + DEPECHE"

# clustering
ptm <- proc.time() # measuring CPU time
pred <-  depeche(tsne_red)
ptm <- proc.time() - ptm

clustering <- pred$clusterVector

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)
number_partitions <- length(table(matched$clusters))

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = matched$labels[labels != "outliers"])
# Add the F1 values to a list 
pops_f1 <- c(pops_f1, cm$byClass[, "F1"])

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, matched$freqs)
# Storing the result 
summ[(nrow(summ)+1),] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                           number_partitions, mf1$found, mf1$w_found, mf1$all, mf1$w_all,
                           tsne_time[1] + tsne_time[4], tsne_time[3], ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
# Print the unmatched contingency matrix:
cat("\n", method, "\n")
print(matched$c)
# Print the matched confusion matrix:
print(cm$table)
```

## ClusterX
```{r tsne clusterx}
method <- "t-SNE + ClusterX"

# clustering
ptm <- proc.time() # measuring CPU time
clustering <- cytof_cluster(tsne_red, method = "ClusterX")
ptm <- proc.time() - ptm

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)
number_partitions <- length(table(matched$clusters))

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = matched$labels[labels != "outliers"])
# Add the F1 values to a list 
pops_f1 <- c(pops_f1, cm$byClass[, "F1"])

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, matched$freqs)
# Storing the result 
summ[(nrow(summ)+1),] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                           number_partitions, mf1$found, mf1$w_found, mf1$all, mf1$w_all,
                           tsne_time[1] + tsne_time[4], tsne_time[3], ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
# Print the unmatched contingency matrix:
cat("\n", method, "\n")
print(matched$c)
# Print the matched confusion matrix:
print(cm$table)
```

# Clustering on UMAP-reduced data (2D)
## RphenoGraph
```{r umap phenograph}
method <- "UMAP + PhenoGraph"

# clustering
ptm <- proc.time() # measuring CPU time
pred <- Rphenograph(umap_red$layout) # default k = 30 nearest neighbours
ptm <- proc.time() - ptm

clustering <- pred$membership

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)
number_partitions <- length(table(matched$clusters))

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = matched$labels[labels != "outliers"])
# Add the F1 values to a list 
pops_f1 <- c(pops_f1, cm$byClass[, "F1"])

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, matched$freqs)
# Storing the result 
summ[(nrow(summ)+1),] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                           number_partitions, mf1$found, mf1$w_found, mf1$all, mf1$w_all,
                           umap_time[1] + umap_time[4], umap_time[3], ptm[1] + ptm[4], ptm[3])
# FOR VISUALIZATION
# Print the unmatched contingency matrix:
cat("\n", method, "\n")
print(matched$c)
# Print the matched confusion matrix:
print(cm$table)
```

## flowMeans
```{r umap flowmeans}
method <- "UMAP + flowMeans"

# clustering
ptm <- proc.time() # measuring CPU time
pred <- flowMeans(umap_red$layout, MaxN = n_pops )
ptm <- proc.time() - ptm

clustering <- pred@Labels[[1]]

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)
number_partitions <- length(table(matched$clusters))

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = matched$labels[labels != "outliers"])
# Add the F1 values to a list 
pops_f1 <- c(pops_f1, cm$byClass[, "F1"])

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, matched$freqs)
# Storing the result 
summ[(nrow(summ)+1),] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                           number_partitions, mf1$found, mf1$w_found, mf1$all, mf1$w_all,
                           umap_time[1] + umap_time[4], umap_time[3], ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
# Print the unmatched contingency matrix:
cat("\n", method, "\n")
print(matched$c)
# Print the matched confusion matrix:
print(cm$table)
```

## flowPeaks
```{r umap flowpeaks}
method <- "UMAP + flowPeaks"

# clustering
ptm <- proc.time() # measuring CPU time
pred <- flowPeaks(umap_red$layout) # default k = 30 nearest neighbours
ptm <- proc.time() - ptm

clustering <- pred$peaks.cluster

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)
number_partitions <- length(table(matched$clusters))

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = matched$labels[labels != "outliers"])
# Add the F1 values to a list 
pops_f1 <- c(pops_f1, cm$byClass[, "F1"])

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, matched$freqs)
# Storing the result 
summ[(nrow(summ)+1),] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                           number_partitions, mf1$found, mf1$w_found, mf1$all, mf1$w_all,
                           umap_time[1] + umap_time[4], umap_time[3], ptm[1] + ptm[4], ptm[3])
# FOR VISUALIZATION
# Print the unmatched contingency matrix:
cat("\n", method, "\n")
print(matched$c)
# Print the matched confusion matrix:
print(cm$table)
```

## DEPECHE
```{r umap depeche}
method <- "UMAP + DEPECHE"

# clustering
ptm <- proc.time() # measuring CPU time
pred <-  depeche(umap_red$layout)
ptm <- proc.time() - ptm

clustering <- pred$clusterVector

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)
number_partitions <- length(table(matched$clusters))

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = matched$labels[labels != "outliers"])
# Add the F1 values to a list 
pops_f1 <- c(pops_f1, cm$byClass[, "F1"])

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, matched$freqs)
# Storing the result 
summ[(nrow(summ)+1),] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                           number_partitions, mf1$found, mf1$w_found, mf1$all, mf1$w_all,
                           umap_time[1] + umap_time[4], umap_time[3], ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
# Print the unmatched contingency matrix:
cat("\n", method, "\n")
print(matched$c)
# Print the matched confusion matrix:
print(cm$table)
```

## ClusterX
```{r umap clusterx}
method <- "UMAP + ClusterX"

# clustering
ptm <- proc.time() # measuring CPU time
clustering <- cytof_cluster(umap_red$layout, method = "ClusterX")
ptm <- proc.time() - ptm

# MATCHING LABELS AND PREDICTIONS
matched <- matching(clustering, labels)
number_partitions <- length(table(matched$clusters))

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                      reference = matched$labels[labels != "outliers"])
# Add the F1 values to a list 
pops_f1 <- c(pops_f1, cm$byClass[, "F1"])

# COMPUTE MEAN F1 and WEIGHTED MEAN F1
mf1 <- mean_f1(cm, matched$freqs)
# Storing the result 
summ[(nrow(summ)+1),] <- c(params$file, nrow(ff@exprs), method, length(table(clustering)),
                           number_partitions, mf1$found, mf1$w_found, mf1$all, mf1$w_all,
                           umap_time[1] + umap_time[4], umap_time[3], ptm[1] + ptm[4], ptm[3])

# FOR VISUALIZATION
# Print the unmatched contingency matrix:
cat("\n", method, "\n")
print(matched$c)
# Print the matched confusion matrix:
print(cm$table)
```

# Results
```{r esults}
kable(summ)
```

















