---
title: "Algorithm benchmarking script"
subtitle: "Synthetic sample test_05"
output:
  html_document:
    number_sections: yes
    theme: journal
    toc: yes
    toc_float: yes
author: Anna Guadall
params:
  file: "ff_05_2019-04-18.fcs"
  labels: "labels_05_2019-04-18"
  freqs: "props_05_2019-04-18"
  seed: 42
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, cache = T)
```

```{r libraries, message=FALSE}
library(FlowSOM)
library(caret)
library(cytofkit)
library(flowPeaks)
library(flowMeans)
library(ggplot2)
library(gridExtra)
library(umap)
library(knitr)
```

# Reading the data
```{r}
# import the fcs file
ff <- flowCore::read.FCS(params$file)

# number of cells
#n <- nrow(ff@exprs)

# Import the labels
labels <- readRDS(params$labels) # "labels" has c levels

# number of populations
c <- length(levels(labels))

# Adding the label "unknown" to the labels:
all_labels <- labels
levels(all_labels) <- c(levels(labels), "unknown") # "all_labels" has c + 1 levels

# Levels list
all_levels <- levels(all_labels)
```

# Random seeds
```{r}
set.seed(params$seed)
seeds <- sample(1:1000, 5, replace = F)
```

# Functions
Matching labels and predictions:
```{r}
matching <- function(prediction, labels, c, all_levels){
  # cross table
  t <- table(prediction, labels)
  
  # sorting max values
  max_values <- apply(t, 2, max)
  sorted_max_values <- sort(max_values)
  
  # Finding the maximum number of each cell type (columns) on each cluster (rows):
  m <- apply(t, 2, which.max)
  
  # Creating a list with n unknowns
  matched_preds <- rep("unknown", length(prediction))
  
  # Replacing the numbers of the clusters by the names of the cell types:
  for(i in 1:length(prediction)){
    for(j in 1:c){ 
      if(prediction[i] == m[names(sorted_max_values)[j]]){ # we compare to the sorted names
        matched_preds[i] <- names(sorted_max_values)[j]
      }
    }
  }

  # Factorize matched predictions including the "unknown" level
  matched_preds <- factor(matched_preds, levels = all_levels)
  
  matched <- list("preds" = matched_preds, "m" = m, "t" = t)
  return(matched)
}
```

In order to compare performances, we will use the F1 score, which is the harmonic average of the precision and recall. The F1 value ranges from 0 (worst performance) to 1 (best performance). One F1 value is computed for every cell type. The mean F1 will be computed for every method in order to give a global performance measure. There are to facts that must be taken into account:

1. The "unknown" class should not be used, as there is not such a class in the original sample.
2. There is no possible to compute the F1 score for the unmatched classes. In this case, the result is "NA". Thus, we are going to replace NAs by zeroes.

Computing the mean F1:
```{r}
mean_f1 <- function(cm, c){
  # Extracting the F1 values
  f1_list <- cm$byClass[,"F1"]
  # replacing NAs by zero
  f1_list[is.na(f1_list)] <-  0
  # Computing mean F1 for all "c" cell labels (without the "unknowns")
  # Storing the result in the environment
  return(mean(f1_list[1:c]))
}
```

# Clustering/partitioning
## FlowSOM
```{r}
# parameters
colors <- grDevices::colorRampPalette(c("pink", "#00007F", "blue", "#007FFF", "cyan", "#7FFF7F", "yellow", "#FF7F00", "red"))
method <- "FlowSOM"

for(i in 1:length(seeds)){
  # meta-clustering
  fs <- FlowSOM(ff, compensate = F,transform = F, scale = F,
                colsToUse = colnames(ff@exprs), nClus = c, seed = seeds[i])
  mc_cell <- fs$metaclustering[fs$FlowSOM$map$mapping[,1]]
  
  # plot
  #PlotStars(fs$FlowSOM, backgroundValues = as.factor(fs$metaclustering))

  # Macth labels and predictions (cell level)
  matched <- matching(mc_cell, labels, c, all_levels)
  # returns 2 elements, matched$preds and matcheds$m
  
  # STORE THE PREDICTIONS (cell level)
  # will I really need this???????????????? maybe for a final graph
  # Generate a name
  name_pred <- paste(method, "pred", i, sep = "_")
  # Assign the name to the prediction
  assign(name_pred, matched$preds)
  
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$preds, reference = all_labels)
  # Print the unmatched cross table:
  print(matched$t)
  # Print the matched confusion matrix:
  print(cm$table)
  # Store the confusion matrix:
  assign(paste(method, "cm", i, sep = "_"), cm)
  
  # COMPUTE THE MEAN F1
  mf1 <- mean_f1(cm, c)
  # Storing the result in the environment
  assign(paste(method, "F1", i, sep = "_"), mf1) 
  
  # Match labels and predictions (node level)
  mc <- rep("NA", length(fs$metaclustering))

  for(i in 1:length(mc)){
    for(j in 1:c){
      if(fs$metaclustering[i] == matched$m[j]){
        mc[i] <- levels(labels)[j]
      }
    }
  }
  
  # Levels of mc in the same order than labels (to take the same color in the plot)
  mc <- factor(mc, levels = levels(labels))
  
  # Plot the results
  PlotPies(fs$FlowSOM, cellTypes = labels, backgroundValues = mc, 
           colorPalette = colors, backgroundColor = colors) 
}
```

## RphenoGraph
```{r}
# clustering
pred <- Rphenograph(ff@exprs) # default k = 30 nearest neighbours

# Macth labels and predictions (cell level)
matched <- matching(pred$membership, labels, c, all_levels)

# Storing the results
pheno_pred <- matched$preds

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
# store and print
print(matched$t)
(pheno_cm <- confusionMatrix(data = matched$preds, reference = all_labels))$table

# COMPUTE THE MEAN F1
pheno_F1 <- mean_f1(pheno_cm, c)
```

## flowMeans
```{r}
pred <- flowMeans(ff@exprs, MaxN = c )

# Macth labels and predictions (cell level)
matched <- matching(pred@Labels[[1]], labels, c, all_levels)

# Storing the results
flowmeans_pred <- matched$preds

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
# store and print
print(matched$t)
(flowmeans_cm <- confusionMatrix(data = matched$preds, reference = all_labels))$table

# COMPUTE THE MEAN F1
flowmeans_F1 <- mean_f1(flowmeans_cm, c)
```

## flowPeaks
```{r}
# clustering
pred <-  flowPeaks(ff@exprs)

# Macth labels and predictions (cell level)
matched <- matching(pred$peaks.cluster, labels, c, all_levels)

# Storing the results
flowpeaks_pred <- matched$preds

# COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
# store and print
print(matched$t)
(flowpeaks_cm <- confusionMatrix(data = matched$preds, reference = all_labels))$table

# COMPUTE THE MEAN F1
flowpeaks_F1 <- mean_f1(flowpeaks_cm, c)
```

## t-SNE
Fixed number of clusters flor flowMeans
```{r}
for(i in 1:length(seeds)){
  #### t-SNE DIMENSIONALITY REDUCTION ####
  # in the example, data is transformed prior to t-SNE reduction
  reduction <-  cytof_dimReduction(data = ff@exprs, method = "tsne", tsneSeed = seeds[i])
  
  # plot:
  p <- ggplot(as.data.frame(reduction), aes(x = tsne_1, y = tsne_2)) + 
        geom_point(aes(color = labels))  + ggtitle("Cell labels") +
        theme(legend.position = "bottom")
  grid.arrange(p, nrow = 1)
  
  #### FLOWMEANS clustering ####
  method <- "tSNE_fMeans"
  pred <- flowMeans(reduction, MaxN = c )
  
  # Macth labels and predictions (cell level)
  matched <- matching(pred@Labels[[1]], labels, c, all_levels)
  
  # STORE THE PREDICTIONS 
  # Generate a name
  name_pred <- paste(method, "pred", i, sep = "_")
  # Assign the name to the prediction
  assign(name_pred, matched$preds)
  
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$preds, reference = all_labels)
  # Print the confusion matrix:
  cat(method, "\n")
  print(matched$t)
  print(cm$table)
  # Store the confusion matrix:
  assign(paste(method, "cm", i, sep = "_"), cm)
  
  # COMPUTE THE MEAN F1
  mf1 <- mean_f1(cm, c)
  # Storing the result in the environment
  assign(paste(method, "F1", i, sep = "_"), mf1) 
  
  
  # plot:
  p <- ggplot(as.data.frame(reduction), aes(x = tsne_1, y = tsne_2)) + 
        geom_point(aes(color = matched$preds)) + ggtitle("flowMeans") +
        theme(legend.position = "bottom")
  grid.arrange(p, nrow = 1)
  

  #### FLOWPEAKS clustering ####
  method <- "tSNE_fPeaks"
  red_data <- as.matrix(reduction)
  colnames(red_data) <- c("tsne_1", "tsne_2")
  pred <-  flowPeaks(red_data)
  
  # Macth labels and predictions (cell level)
  matched <- matching(pred$peaks.cluster, labels, c, all_levels)
  
  # STORE THE PREDICTIONS 
  # Generate a name
  name_pred <- paste(method, "pred", i, sep = "_")
  # Assign the name to the prediction
  assign(name_pred, matched$preds)
  
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$preds, reference = all_labels)
  # Print the confusion matrix:
  cat(method, "\n")
  print(matched$t)
  print(cm$table)
  # Store the confusion matrix:
  assign(paste(method, "cm", i, sep = "_"), cm)
  
  # COMPUTE THE MEAN F1
  mf1 <- mean_f1(cm, c)
  # Storing the result in the environment
  assign(paste(method, "F1", i, sep = "_"), mf1) 
  
  # plot
  p <- ggplot(as.data.frame(reduction), aes(x = tsne_1, y = tsne_2)) + 
        geom_point(aes(color = matched$preds)) + ggtitle("flowPeaks") +
        theme(legend.position = "bottom")
  grid.arrange(p, nrow = 1)
}
```  
  
## UMAP
```{r}
for(i in 1:length(seeds)){
  #### UMAP DIMENSIONALITY REDUCTION ####
  reduction <- umap(ff@exprs, random_state = seeds[i])
  colnames(reduction$layout) <- c("UMAP_1", "UMAP_2")

  # plot:
  p <- ggplot(as.data.frame(reduction$layout), aes(x = UMAP_1, y = UMAP_2)) + 
        geom_point(aes(color = labels))  + ggtitle("Cell labels") +
        theme(legend.position = "bottom")
  grid.arrange(p, nrow = 1)
  
  #### FLOWMEANS clustering ####
  method <- "UMAP_fMeans"
  pred <- flowMeans(reduction$layout, MaxN = c )
  
  # Macth labels and predictions (cell level)
  matched <- matching(pred@Labels[[1]], labels, c, all_levels)
  
  # STORE THE PREDICTIONS 
  # Generate a name
  name_pred <- paste(method, "pred", i, sep = "_")
  # Assign the name to the prediction
  assign(name_pred, matched$preds)
  
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$preds, reference = all_labels)
  # Print the confusion matrix:
  cat(method, "\n")
  print(matched$t)
  print(cm$table)
  # Store the confusion matrix:
  assign(paste(method, "cm", i, sep = "_"), cm)
  
  # COMPUTE THE MEAN F1
  mf1 <- mean_f1(cm, c)
  # Storing the result in the environment
  assign(paste(method, "F1", i, sep = "_"), mf1) 
  
  # plot:
  p <- ggplot(as.data.frame(reduction$layout), aes(x = UMAP_1, y = UMAP_2)) + 
        geom_point(aes(color = matched$preds)) + ggtitle("flowMeans") +
        theme(legend.position = "bottom")
  grid.arrange(p, nrow = 1)
  

  #### FLOWPEAKS clustering ####
  method <- "UMAP_fPeaks"
  red_data <- as.matrix(reduction$layout)
  colnames(red_data) <- c("tsne_1", "tsne_2")
  pred <-  flowPeaks(red_data)
  
  # Macth labels and predictions (cell level)
  matched <- matching(pred$peaks.cluster, labels, c, all_levels)
  
  # STORE THE PREDICTIONS 
  # Generate a name
  name_pred <- paste(method, "pred", i, sep = "_")
  # Assign the name to the prediction
  assign(name_pred, matched$preds)
  
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$preds, reference = all_labels)
  # Print the confusion matrix:
  cat(method, "\n")
  print(matched$t)
  print(cm$table)
  # Store the confusion matrix:
  assign(paste(method, "cm", i, sep = "_"), cm)
  
  # COMPUTE THE MEAN F1
  mf1 <- mean_f1(cm, c)
  # Storing the result in the environment
  assign(paste(method, "F1", i, sep = "_"), mf1) 
  
  # plot
  p <- ggplot(as.data.frame(reduction$layout), aes(x = UMAP_1, y = UMAP_2)) + 
        geom_point(aes(color =  matched$preds)) + ggtitle("flowPeaks") +
        theme(legend.position = "bottom")
  grid.arrange(p, nrow = 1)
}
```  

# Evaluation
```{r}
# t-SNE to show the results:
reduction <-  cytof_dimReduction(data = ff@exprs, method = "tsne", tsneSeed = sum(seeds)) # different seed

# Information about pop's frequencies
percent <-  readRDS(params$freqs)
```

## RPhenograph, flowMeans, flowPeaks
```{r}
# mean F1 table
summ1 <- data.frame(Mean_F1 = 0)
summ1[1,] <- round(pheno_F1, 3)
summ1[2,] <- round(flowmeans_F1, 3)
summ1[3,] <- round(flowpeaks_F1, 3)

rownames(summ1) <- c("RPhenograph", "flowMeans", "flowPeaks")
kable(summ1)
```

```{r}
# Plot all F1 results
class <- factor(all_levels[1:c], levels = levels(labels))
method <- c("RPhenograph", "flowMeans", "flowPeaks")
method <- factor(method, levels = method)
df <- data.frame(class = rep(class,3), method = rep(method, each = c), 
                 F1 = c(pheno_cm$byClass[1:c,"F1"], flowmeans_cm$byClass[1:c,"F1"], flowpeaks_cm$byClass[1:c,"F1"]),
                 percent = rep(percent,3))
ggplot(df, aes(class, F1)) + geom_point(aes(size = percent, color = percent)) + 
  coord_flip() + ggtitle("RPhenograph") + ylim(0,1) + facet_wrap(.~method)

```

```{r}
# preparing data for plotting:
data <- cbind(as.data.frame(reduction), as.data.frame(labels), 
              as.data.frame(pheno_pred), as.data.frame(flowmeans_pred), as.data.frame(flowpeaks_pred))
```

```{r, fig.width=7, fig.height=8}
p1 <- cytof_clusterPlot(data = data, xlab = "tsne_1", ylab = "tsne_2", 
                  cluster = "labels", sampleLabel = FALSE, labelSize = 4, title = "Cell labels")
p2 <- cytof_clusterPlot(data = data, xlab = "tsne_1", ylab = "tsne_2", 
                  cluster = "pheno_pred", sampleLabel = FALSE, labelSize = 4, title = "RPhenograph")
p3 <- cytof_clusterPlot(data = data, xlab = "tsne_1", ylab = "tsne_2", 
                  cluster = "flowmeans_pred", sampleLabel = FALSE, labelSize = 4, title = "flowMeans")
p4 <- cytof_clusterPlot(data = data, xlab = "tsne_1", ylab = "tsne_2",
                  cluster = "flowpeaks_pred", sampleLabel = FALSE, labelSize = 4, title = "flowPeaks")

grid.arrange(p1, p2, p3, p4, nrow = 2)
```

## FlowSOM, t-SNE, UMAP
```{r}
summ2 <- data.frame(Seed_1 = 0, Seed_2 = 0, Seed_3 = 0, Seed_4 = 0, Seed_5 = 0)

summ2[1,] <- round( c(FlowSOM_F1_1, FlowSOM_F1_2, FlowSOM_F1_3, FlowSOM_F1_4, FlowSOM_F1_5), 3)
summ2[2,] <- round( c(tSNE_fMeans_F1_1, tSNE_fMeans_F1_2, tSNE_fMeans_F1_3, tSNE_fMeans_F1_4, tSNE_fMeans_F1_5), 3)
summ2[3,] <- round( c(tSNE_fPeaks_F1_1, tSNE_fPeaks_F1_2, tSNE_fPeaks_F1_3, tSNE_fPeaks_F1_4, tSNE_fPeaks_F1_5), 3)
summ2[4,] <- round( c(UMAP_fMeans_F1_1, UMAP_fMeans_F1_2, UMAP_fMeans_F1_3, UMAP_fMeans_F1_4, UMAP_fMeans_F1_5), 3)
summ2[5,] <- round( c(UMAP_fPeaks_F1_1, UMAP_fPeaks_F1_2, UMAP_fPeaks_F1_3, UMAP_fPeaks_F1_4, UMAP_fPeaks_F1_5), 3)
summ2$average_mean_F1 <- round(apply(summ2, 1, mean),3) # compute mean
summ2$SD_F1 <- round(apply(summ2[,-ncol(summ2)], 1, sd),3) # compute SD

rownames(summ2) <- c("FlowSOM", "t-SNE + flowMeans", "t-SNE + flowPeaks", "UMAP + flowMeans", "UMAP + flowPeaks")
kable(summ2)
```

```{r}
# Plot all F1 results

# preparing the data
class <- factor(all_levels[1:c], levels = levels(labels))
method <-  c("FlowSOM", "t-SNE + flowMeans", "t-SNE + flowPeaks", "UMAP + flowMeans", "UMAP + flowPeaks")
method <- factor(method, levels = method)
conditions <- length(method) * length(seeds)

# confusion matrices
cm_list <- c("FlowSOM_cm", "tSNE_fMeans_cm", "tSNE_fPeaks_cm", "UMAP_fMeans_cm", "UMAP_fPeaks_cm")

F1 <- c()
for(i in cm_list){
  j <- ls(pattern = i)
  for(k in 1:length(j)){
    for(l in 1:c){
      F1 <-  c(F1, eval(parse(text=j[k]))$byClass[l,"F1"])
    }
  }
}

# preparing data frame
df2 <- data.frame(class = rep(class, conditions), method = rep(method, each = c*length(seeds)), 
                 F1 = F1, percent = rep(percent, conditions))

# plotting
ggplot(df2, aes(class, F1)) + geom_boxplot(aes(color = percent)) + 
  coord_flip() + ggtitle("F1") + ylim(0,1) + facet_wrap(.~method)

```

```{r}
# plot all predictions

# matched predictions
pred_list <- c("FlowSOM_pred", "tSNE_fMeans_pred", "tSNE_fPeaks_pred", "UMAP_fMeans_pred", "UMAP_fPeaks_pred")

predictions <- c()
method_i <- c()
for(i in pred_list){
  j <- ls(pattern = i)
  for(k in 1:length(j)){
    predictions <-  c(predictions, eval(parse(text=j[k])))
    method_i <- c(method_i, j[k])
  }
}

# preparing data frame
df3 <- data.frame(tsne_1 = rep(reduction[,1], length(pred_list)*length(seeds)), 
                  tsne_2 = rep(reduction[,2], length(pred_list)*length(seeds)),
                  method = rep(method_i, each = length(labels)), 
                  predictions = predictions)
```

```{r, fig.width=10, fig.height=15}
# plotting
ggplot(df3, aes(x = tsne_1, y = tsne_2)) + geom_point(aes(color = factor(predictions))) + 
  facet_wrap(.~method)
```


```{r, eval = F, cache = F}
# export results
write.table(cbind(params$file, rownames(summ1), summ1), 
            file = "../results_f1/oneseed.txt", append = T, sep = "\t", row.names = F, col.names = F)
write.table(cbind(params$file, rownames(summ2), summ2), 
            file = "../results_f1/fiveseeds.txt", append = T, sep = "\t", row.names = F, col.names = F)
```



