---
title: "Installing packages for cytometry analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

http://bioconductor.org/install/

# Bioconductor
## BiocManager
Necessary to install Bioconductor Packages
```{r}
if (!requireNamespace("BiocManager"))
    install.packages("BiocManager")
BiocManager::install()
```

## Bioconductor Packages
### BiocStyle
```{r}
BiocManager::install(c("BiocStyle"))
```

### flowCore
Basic Functions for Flow Cytometry Data
```{r}
BiocManager::install("flowCore", version = "3.8")
```
```{r}
browseVignettes("flowCore")
```
The pdf on vignettes has date January 4, 2019. This vignette has several errors. Find the last pdf of February 26, 2019 here: https://www.bioconductor.org/packages/devel/bioc/vignettes/flowCore/inst/doc/HowTo-flowCore.pdf

### ggcyto
A ggplot2 graphics implementation.
```{r}
BiocManager::install("ggcyto", version = "3.8")
```
```{r}
browseVignettes("ggcyto")
```

### flowWorkspaceData
```{r}
BiocManager::install("flowWorkspaceData", version = "3.8")
```

### SPADE
package ‘spade’ is not available (for R version 3.5.2)  
package ‘spade’ has been removed with Bioconductor 3.4 release  
This package is for version 2.12 of Bioconductor  

### FlowSOM
FlowSOM offers visualization options for cytometry data, by using Self-Organizing Map clustering and Minimal Spanning Trees.
```{r}
BiocManager::install("FlowSOM", version = "3.8")
```
```{r}
browseVignettes("FlowSOM")
```

### oneSENSE
```{r}
BiocManager::install("oneSENSE", version = "3.8")
```
```{r}
browseVignettes("oneSENSE")
```

### flowMAP
Mapping cell populations in flow cytometry data for cross-sample comparisons using the Friedman-Rafsky Test  
Friedman-Rafskty (FR) statistic to compare similarity between cell populations across multiple flow cytometry samples  
determine cell population groups across samples
```{r}
BiocManager::install("flowMap", version = "3.8")
```
```{r}
browseVignettes("flowMap")
```

### CytoML
A GatingML Interface for Cross Platform Cytometry Data Sharing
```{r}
BiocManager::install("CytoML", version = "3.8")
```

### flowAI
Automatic and GUI methods to do quality control on Flow cytometry Data
```{r}
BiocManager::install("flowAI", version = "3.8")
```
```{r}
browseVignettes("flowAI")
```

### flowClean
performing quality control on flow cytomery datasets.
```{r}
BiocManager::install("flowClean", version = "3.8")
```
```{r}
browseVignettes("flowClean")
```

### flowMeans
Non-parametric Flow Cytometry Data Gating
```{r}
BiocManager::install("flowMeans", version = "3.8")
```
```{r}
browseVignettes("flowMeans")
```

### flowPeaks
```{r}
BiocManager::install("flowPeaks", version = "3.8")
```
```{r}
browseVignettes("flowPeaks")
```

### SamSPECTRAL
```{r}
BiocManager::install("SamSPECTRAL", version = "3.8")
```
```{r}
browseVignettes("SamSPECTRAL")
```

### Cytofast
Quickly visualizing and analyzing the output of Cytosplore; software used to characterizes and cluster immune subsets from flow or mass cytometry (CyTOF) data.  
IMAGING FOR CLUSTERS  
http://www.bioconductor.org/packages/devel/bioc//vignettes/cytofast/inst/doc/spitzer.html  
https://www.sciencedirect.com/science/article/pii/S200103701830059X
```{r}
BiocManager::install("cytofast")
```

## Updating installed Bioconductor Packages
```{r}
BiocManager::install()
```
```{r}
browseVignettes("CytoML")
```


# Not in Bioconductor
## UMAP
Uniform manifold approximation and projection is a technique for dimension reduction.   
https://cran.r-project.org/web/packages/umap/index.html
```{r}
install.packages("umap")
```
https://cran.r-project.org/web/packages/umap/vignettes/umap.html  

## Rphenograph
```{r}
if(!require(devtools)){
  install.packages("devtools") # If not already installed
}
devtools::install_github("JinmiaoChenLab/Rphenograph")
```
https://github.com/JinmiaoChenLab/Rphenograph

## Cytofkit
cytofkit: an integrated mass cytometry data analysis pipeline  
https://www.rdocumentation.org/packages/cytofkit/versions/1.4.8
```{r}
devtools::install_github("JinmiaoChenLab/cytofkit")
```

## Rclusterpp
clang: error: unsupported option '-fopenmp'

Large-scale hierarchical clustering in R  
https://github.com/nolanlab/Rclusterpp  
https://github.com/nolanlab/Rclusterpp/wiki/Getting-Started  
Installation using MacPorts and g++4.8 (With OpenMP Support)  

I install XCode 10.1  
https://itunes.apple.com/us/app/xcode/id497799835?ls=1&mt=12  

I install MacPorts  
https://www.macports.org/install.php  
After installation is done, it is recommended that you run this step manually on a regular basis to to keep your MacPorts system always current:  
sudo port -v selfupdate  

I try to install gcc version 4.8: sudo port install gcc48  
Error: building gcc48 is not supported with Xcode 9 or greater
Error: Failed to fetch gcc48: unsupported platform

In fact:
gcc --version
Configured with: --prefix=/Applications/Xcode.app/Contents/Developer/usr --with-gxx-include-dir=/Applications/Xcode.app/Contents/Developer/Platforms/MacOSX.platform/Developer/SDKs/MacOSX10.14.sdk/usr/include/c++/4.2.1
Apple LLVM version 10.0.0 (clang-1000.11.45.5)
Target: x86_64-apple-darwin18.2.0
Thread model: posix
InstalledDir: /Applications/Xcode.app/Contents/Developer/Toolchains/XcodeDefault.xctoolchain/usr/bin

I install /Library/Developer/CommandLineTools/Packages/macOS_SDK_headers_for_macOS_10.14.pkg 

**THIS HAS BEEN THE SOLUTION:**
https://github.com/velocyto-team/velocyto.R/issues/2
Install gfortran as suggested: https://gcc.gnu.org/wiki/GFortranBinaries  
Homebrew offers the latest stable gfortran release as part of the "gcc" package.
First, I install Homebrew  
Then: brew install gcc  

Download clang-4.0.0-darwin15.6-Release.tar.gz and uncompress.  
Move to /usr/local/  

create ~/.R/Makevars such as
CC=/usr/local/clang4/bin/clang
CXX=/usr/local/clang4/bin/clang++
LDFLAGS=-L/usr/local/clang4/lib

export PATH=/usr/local/clang4/bin:$PATH
**Should I add it to ~/.bash_profile?**

```{r}
devtools::install_github("nolanlab/Rclusterpp")
```

## SPADE
Needs Rclusterpp
```{r}
devtools::install_github("nolanlab/spade")
```


## scaffold
https://github.com/ParkerICI/flow-analysis-tutorial  
Scaffold has been divided in grappolo, vite and panorama.  

### grappolo
clustering of .fcs files  
https://github.com/ParkerICI/grappolo
```{r}
devtools::install_github("ParkerICI/grappolo")
```

### vite
This package enables the analysis of single-cell data using graphs, both unsupervised graphs as well as scaffold maps. While the package is designed to work with clusters generated from the grappolo package, any kind of tabular input data can be used as input.  
https://github.com/ParkerICI/vite
```{r}
devtools::install_github("ParkerICI/vite")
```

### panorama
Visualization of the graph
https://github.com/ParkerICI/panorama
```{r}
devtools::install_github("ParkerICI/panorama")
```


## Citrus
https://github.com/nolanlab/citrus/wiki/Installing-Citrus
```{r}
devtools::install_github('nolanlab/citrus')
```
Error: package or namespace load failed for ‘samr’ in loadNamespace(j <- i[[1L]], c(lib.loc, .libPaths()), versionCheck = vI[[j]]):
 there is no package called ‘impute’
Error : package ‘samr’ could not be loaded
ERROR: lazy loading failed for package ‘citrus’

```{r}
BiocManager::install("impute", version = "3.8")
```
With impute I can install citrus.


# Not in R
## Vortex
I download the latest release: https://github.com/nolanlab/vortex/releases  
System requirements: Java 1.8 64-bit  
java -version :  
java version "1.6.0_65"  
Java(TM) SE Runtime Environment (build 1.6.0_65-b14-468)  
Java HotSpot(TM) 64-Bit Server VM (build 20.65-b04-468, mixed mode)  

I install jenv from brew (to manage java environments):  
https://stackoverflow.com/questions/33259758/switching-from-java-1-6-to-1-8-on-a-mac  
brew install jenv  
echo 'export PATH="$HOME/.jenv/bin:$PATH"' >> ~/.bash_profile  
echo 'eval "$(jenv init -)"' >> ~/.bash_profile  
exec $SHELL -l





