---
title: "real_02 downsample 2 (temoin 1)"
output: html_document
params:
  experience: "real_02"
  file: "real_02.wsp"
  sample: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(flowWorkspace)
library(caret)
```

# Importing data from FlowJo workspace
Open FlowJo workspace
```{r}
(ws <-  openWorkspace(params$file))
```

# Parsing the workspace: GATING SET
```{r}
gs <-  parseWorkspace(ws, name = 2, sampNloc="sampleNode") # 1: All Samples, 2: group
```

```{r}
gs@data
```

Gating strategy:
```{r}
plot(gs)
```

# Extracting the data from one sample
The sample is defined in the parameters:
```{r}
(s <- params$sample)
```

```{r}
getSamples(ws)[s,]
```

```{r}
getSamples(ws)[s,"name"]
```

Data:
```{r}
gs@data[[s]]
```

```{r}
summary(gs@data[[s]])
```

Counts:
```{r}
getPopStats(gs[[s]])
```

Variation between `FlowJo` and `openCyto` counts:
```{r}
plotPopCV(gs[[s]])
```

(The `openCyto`package is included in `flowWorkspace`).

Extracting the data
```{r}
data <- exprs(gs@data[[s]])
# Or, equivalently:
# data  <- getData(gs[[s]])
nrow(data)
```

## Extracting cell labels
Preparing the data frame
```{r}
head(data)
```

We are just interested in columns 7-11
```{r}
data <- data[,7:11]
data <- as.data.frame(data)
head(data)
```

Getting the nodes (or the populations):
```{r}
(nodes <- getNodes(gs, path = "auto"))
```

We can extract the cells labels (indices) resulting from applying the gates imported from `FlowJo` (there might be slight differences). 
For example:
```{r}
nkt <- getIndices(gs[[s]], nodes[6])
table(nkt)
```


Let's first see which cell populations we want to label:

```{r}
getNodes(gs)
```

## CELLS
We are just interested in viable, single cells lymphocytes (no monocytes): node 5.
```{r}
nodes[5]
cells <- getIndices(gs[[s]], nodes[5])
length(cells)
table(cells)
```

Let's tag them as "lymphocytes". The rest, will be "other_events".
```{r}
data$cells <-  "other_events" 
data$cells <- as.factor(ifelse(cells, "lymphocytes", data$cells))
data$cells <- ifelse(cells, "lymphocytes", data$cells)
table(data$cells)
```

## LABELS
This will be the SUBPOPULATIONS.  

All the ungated cells will be tagged as "outliers".
```{r}
data$labels <- "outliers"
```

Naming the labels:
```{r}
label_names <- c("7" = "NKT_4", "8" = "NKT_DP", "9" = "NKT_8", "10" = "NKT_DN",
                 "12" = "NK_4", "13" = "NK_DP", "14" = "NK_8", "15" = "NK_DN",
                 "17" = "T_cells_4", "18" = "T_cells_DP", "19" = "T_cells_8", "20" = "T_cells_DN",
                 "23" = "B_4", "24" = "B_DP", "25" = "B_8", "26" = "B_DN",
                 "28" = "NO_BTNK_4", "29" = "NO_BTNK_DP", "30" = "NO_BTNK_8", "31" = "NO_BTNK_DN")
names(label_names)
```

Labelling:
```{r}
for(i in names(label_names)){
    indices <- getIndices(gs[[s]], nodes[as.numeric(i)])
    # if index for node i is TRUE, data$label takes label_name[i]
    # otherwise, it remains unchanged
    data$labels <- ifelse(indices, label_names[[i]], data$labels)
}

data$labels <-  as.factor(data$labels)
table(data$labels)
```

## META-LABELS
We can now group the labels. For instance, we will follow the gating logic:
```{r}
# Empty column
data$metalabels <- "outliers"

# Naming the meta-labels and the patterns that will be used to define them
metalabels_list <- c("NKT", "NK", "T", "B", "NO_BTNK")
patterns_list <- c("NKT_", "NK_", "T_cells_", "B_", "NO_B")

for(i in 1:length(metalabels_list)){
  data$metalabels <- ifelse(grepl(patterns_list[i], data$labels), metalabels_list[i], data$metalabels)
}

data$metalabels <- as.factor(data$metalabels)
table(data$metalabels)
```

```{r}
table(data$labels, data$metalabels)
```

## LYMPHOCYTES
We are going to keep just the living, single cell lymphocytes:
```{r}
data <- data[data$cells == "lymphocytes",]
table(data$labels, data$metalabels)
```


# Export data 
## Conversion to `FlowFrame` class objects
```{r}
head(data)
```

Replacing the fluorophores by the markers:
```{r}
colnames(data)
```

```{r}
colnames(data)[1:5] <-  c("CD3", "CD4", "CD8", "CD56", "CD19")
colnames(data)
```

##Downsampling
Based on "labels".
```{r}
subsample <- createDataPartition(data$labels, p = 0.15, list = F)
```

```{r}
library(flowCore)

# Just the live, single-cell lymphocytes!
#ff <- new("flowFrame", exprs = as.matrix(data[data$cells == "lymphocytes", 1:5])) # 1:5 markers, w/o the labels

ff <- new("flowFrame", exprs = as.matrix(data[subsample, 1:5])) # 1:5 markers, w/o the labels
summary(ff)
```

```{r}
library(ggcyto)
autoplot(ff[], colnames(data)[4], colnames(data)[1])
```

## Export to .fcs
```{r, eval = T}
#write.FCS(ff, paste("ff_", params$experience, "_", getSamples(ws)[s,"name"], ".fcs", sep = ""))
write.FCS(ff, paste("subsample_ff_", params$experience, "_", getSamples(ws)[s,"name"], ".fcs", sep = ""))
```

## Exporting the sublabels groups:
```{r}
sublabels_levels <- list("NKT" = c("NKT_4", "NKT_DP", "NKT_8", "NKT_DN"),
                         "NK" = c("NK_4", "NK_DP", "NK_8", "NK_DN"),
                         "T" = c("T_cells_4", "T_cells_DP", "T_cells_8", "T_cells_DN"),
                         "B" = c("B_4", "B_DP", "B_8", "B_DN"),
                         "NO_BTNK" = c("NO_BTNK_4", "NO_BTNK_DP", "NO_BTNK_8", "NO_BTNK_DN"))
saveRDS(sublabels_levels, paste("sublabels_levels", params$experience, sep = "_"))
```

## Save the labels and meta-labels
```{r}
data$metalabels <- factor(data$metalabels, levels = c("NKT", "NK", "T", "B", "NO_BTNK", "outliers"))
```

```{r, eval = F}
saveRDS(data[, "labels"],
        paste("labels_", params$experience, "_", getSamples(ws)[s,"name"], sep = "")) 
saveRDS(data[, "metalabels"],
        paste("metalabels_", params$experience, "_", getSamples(ws)[s,"name"], sep = "")) 
```

```{r, eval = T, cache = F}
saveRDS(data[subsample, "labels"],
        paste("subsample_labels_", params$experience, "_", getSamples(ws)[s,"name"], sep = "")) 
saveRDS(data[subsample, "metalabels"],
        paste("subsample_metalabels_", params$experience, "_", getSamples(ws)[s,"name"], sep = "")) 
```

```{r}
closeWorkspace(ws)
```


