---
title: "Preparing semi-synthetic data: 11 parameters"
output: html_document
params:
  experience: "20190716_billes"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = T)
```

```{r}
library(flowWorkspace)
library(caret)
library(flowCore)
```

# Introduction
An `R Markdown` script has been developed in order to prepare flow cytometry data for algorithm benchmarking. The data has been previously pre-treated and gated in `FlowJo`. This script enables importing and exploring the data, applying the gating strategy performed on `FlowJo`, labelling the examples according to this gating, performing down-sampling and exporting the data for further analysis. 

# Data
Stained beads (11 markers), 18 detectors.

# Parameters
```{r}
params
```

# Open `fcs` files
## Compensated
```{r}
comp <- list.files(".", pattern = "compens_18") # Compensated
```

```{r}
colnames(read.FCS("compens_18_01_NK.fcs")@exprs)
```
\  

We are interested in columns 1:22.

Files:
```{r}
comp
```
\  

Label list:
```{r}
label_list <- c("P01", "P02", "P03", "P04", "P05", "P06", "P07", "P08", "P09", "P10", 
                "P11", "P12", "P13", "P14", "P15", "P16", "P17", "P18", "P19", "P20")
```

Open the compensated files and put the selected columns in a labelled dataframe,:
```{r}
j <- 1
for(i in 1:length(comp)){
  c <- as.data.frame(read.FCS(comp[i])@exprs[,1:22])
  c <- cbind(c, label = label_list[i])
  if(j == 1){
    data_comp <- c
  }else{
    data_comp <- rbind(data_comp, c)
  }
  j <- j + 1
  rm(c)
}

summary(data_comp[,"label"])
```

## Uncompensated data
```{r}
( uncomp <- list.files(".", pattern = "uncomp") )
```

Open the uncompensated files and put the selected columns in a labelled dataframe,:
```{r}
j <- 1
for(i in 1:length(uncomp)){
  c <- as.data.frame(read.FCS(uncomp[i])@exprs[,1:22])
  c <- cbind(c, label = label_list[i])
  if(j == 1){
    data_uncomp <- c
  }else{
    data_uncomp <- rbind(data_uncomp, c)
  }
  j <- j + 1
  rm(c)
}

table(summary(data_comp[,"label"]) == summary(data_uncomp[,"label"]))
```

# Naming the variables
```{r}
colnames(data_comp)
```

Replacing the fluorophores by the markers:
```{r}
colnames(data_comp)[5:22] <-  c("355-A-A", "355-B-A", "CD25_BV786",
                           "405-B-A", "CD45RA_BV650", "405-D-A",
                           "405-E-A", "CD27_BV786", "CD8a_PerCP-Cy55", 
                           "CD4_BB515", "CCR7_PE-Cy7", "561-B-A",
                           "561-C-A", "CD19_PECF554", "CD3_PE",
                           "CD56_APCH7", "CD127_APC-R700", "CD38-APC")
colnames(data_uncomp) <- colnames(data_comp)
colnames(data_comp)
```

# Frequencies
Bead frequencies are completely arbitrary.
```{r}
(comp_freq <- (prop.table(table(data_comp$label))))
```

## Equivalent frequencies:
```{r}
# looking for the minimum
min_beads <-  which.min(table(data_comp$label))

table(data_comp$label)[min_beads]
```

We could do 15621*20 = 312420 beads, but it is weird.  
We wil take 15000 beads/pop in order to have a final sample with 300000 beads.

```{r}
n <- 15000
down_sample <- c()

set.seed(23)
for(i in 1:length(label_list)){
  sel <- sample(rownames(data_comp[data_comp$label == label_list[i],]), n, replace = F) 
  down_sample <- c(down_sample, sel)
  rm(sel)
}

data_comp_eq <- data_comp[down_sample,]
table(data_comp_eq$label)
```

Great
```{r}
data_uncomp_eq <- data_uncomp[down_sample,]
table(data_uncomp_eq$label)
```

## Different frequencies
Let's use the frequancies used to generate synthetic samples for this experimental design:
```{r}
(theo_props <- readRDS("props_13_a_2019-08-01")/100 )
```

Which is the limitant?  
If we take the max freq:
```{r}
max_freq <- which.max(theo_props)
theo_props[max_freq]
table(data_comp$label)[max_freq]
```

So, the total number of beads will be:
```{r}
(tot <- floor(table(data_comp$label)[max_freq]/theo_props[max_freq]))
```

Thus, the number of cells for each population will be:
```{r}
(ns <- round(tot * theo_props))
```

```{r}
sum(ns)
```

If we substract these numbers to the toal number of beads:
```{r}
table(data_comp$label) - ns
```

Great. Let's do the sample:
```{r}
down_sample <- c()

set.seed(23)
for(i in 1:length(theo_props)){
  sel <- sample(rownames(data_comp[data_comp$label == label_list[i],]), ns[i], replace = F) 
  down_sample <- c(down_sample, sel)
  rm(sel)
}

data_comp_diff <- data_comp[down_sample,]
table(data_comp_diff$label)
```

```{r}
data_uncomp_diff <- data_comp[down_sample,]
table(table(data_uncomp_diff$label) == table(data_comp_diff$label))
```


# Rearranging
We have 3 kind of samples, in compensate and uncompensate version.  
Let's randomly rearrange all of them:
```{r}
## ALL THE BEADS
(n <- nrow(data_comp))
```

```{r}
# Randomly sampling
set.seed(42)
x <- sample(1:n, n, replace = F)

# Rearranging the dataframes
data_comp <-data_comp[x,] 
data_uncomp <-data_uncomp[x,] 

# Renaming the rows (in order)
rownames(data_comp) <- 1:n 
rownames(data_uncomp) <- 1:n 

## EQUAL FREQUENCIES
n <- nrow(data_comp_eq)

# Randomly sampling
set.seed(42)
x <- sample(1:n, n, replace = F)

# Rearranging the dataframes
data_comp_eq <-data_comp_eq[x,] 
data_uncomp_eq <-data_uncomp_eq[x,] 

# Renaming the rows (in order)
rownames(data_comp_eq) <- 1:n 
rownames(data_uncomp_eq) <- 1:n 

## DIFFERENT FREQUENCIES
n <- nrow(data_comp_diff)

# Randomly sampling
set.seed(42)
x <- sample(1:n, n, replace = F)

# Rearranging the dataframes
data_comp_diff <-data_comp_diff[x,] 
data_uncomp_diff <-data_uncomp_diff[x,] 

# Renaming the rows (in order)
rownames(data_comp_diff) <- 1:n 
rownames(data_uncomp_diff) <- 1:n 
```


# Conversion to `FlowFrame` class objects and export to `.fcs`
## All parameters
```{r}
library(flowCore)
ff_comp <- new("flowFrame", exprs = as.matrix(data_comp[, 1:22])) # 1:22 channels, w/o the labels
ff_uncomp <- new("flowFrame", exprs = as.matrix(data_uncomp[, 1:22]))
ff_comp_eq <- new("flowFrame", exprs = as.matrix(data_comp_eq[, 1:22])) 
ff_uncomp_eq <- new("flowFrame", exprs = as.matrix(data_uncomp_eq[, 1:22]))
ff_comp_diff <- new("flowFrame", exprs = as.matrix(data_comp_diff[, 1:22]))  
ff_uncomp_diff <- new("flowFrame", exprs = as.matrix(data_uncomp_diff[, 1:22]))
summary(ff_comp)
```

```{r}
write.FCS(ff_comp, paste("ff_comp_18_all_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_uncomp, paste("ff_non_comp_18_all_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_comp_eq, paste("ff_comp_eq_18_all_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_uncomp_eq, paste("ff_non_comp_18_eq_all_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_comp_diff, paste("ff_comp_diff_18_all_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_uncomp_diff, paste("ff_non_comp_diff_18_all_", params$experience, "_", params$beads, ".fcs", sep = ""))
```

## All detectors, without FSC, SSC
```{r}
ff_comp <- new("flowFrame", exprs = as.matrix(data_comp[, 5:22])) # 1:22 channels, w/o the labels
ff_uncomp <- new("flowFrame", exprs = as.matrix(data_uncomp[, 5:22]))
ff_comp_eq <- new("flowFrame", exprs = as.matrix(data_comp_eq[, 5:22])) 
ff_uncomp_eq <- new("flowFrame", exprs = as.matrix(data_uncomp_eq[, 5:22]))
ff_comp_diff <- new("flowFrame", exprs = as.matrix(data_comp_diff[, 5:22]))  
ff_uncomp_diff <- new("flowFrame", exprs = as.matrix(data_uncomp_diff[, 5:22]))
summary(ff_comp)
```

```{r}
write.FCS(ff_comp, paste("ff_comp_18_fluo_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_uncomp, paste("ff_non_comp_18_fluo_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_comp_eq, paste("ff_comp_eq_18_fluo_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_uncomp_eq, paste("ff_non_comp_eq_18_fluo_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_comp_diff, paste("ff_comp_diff_18_fluo_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_uncomp_diff, paste("ff_non_comp_diff_18_fluo_", params$experience, "_", params$beads, ".fcs", sep = ""))
```

## 11 detectors (with FSC, SSC)
```{r}
colnames(data_comp)
```

```{r}
(c <- colnames(data_comp)[c(1,2,3,4,7,9,12,13,14,15,18,19,20,21,22)])
```

```{r}
ff_comp <- new("flowFrame", exprs = as.matrix(data_comp[, c])) # 1:22 channels, w/o the labels
ff_uncomp <- new("flowFrame", exprs = as.matrix(data_uncomp[, c]))
ff_comp_eq <- new("flowFrame", exprs = as.matrix(data_comp_eq[, c])) 
ff_uncomp_eq <- new("flowFrame", exprs = as.matrix(data_uncomp_eq[, c]))
ff_comp_diff <- new("flowFrame", exprs = as.matrix(data_comp_diff[, c]))  
ff_uncomp_diff <- new("flowFrame", exprs = as.matrix(data_uncomp_diff[, c]))
summary(ff_comp)
```

```{r}
write.FCS(ff_comp, paste("ff_comp_11_all_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_uncomp, paste("ff_non_comp_11_all_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_comp_eq, paste("ff_comp_eq_11_all_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_uncomp_eq, paste("ff_non_comp_11_eq_all_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_comp_diff, paste("ff_comp_diff_11_all_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_uncomp_diff, paste("ff_non_comp_diff_11_all_", params$experience, "_", params$beads, ".fcs", sep = ""))
```

## 11 detectors (without FSC, SSC)
```{r}
(c <- colnames(data_comp)[c(7,9,12,13,14,15,18,19,20,21,22)])
```

```{r}
ff_comp <- new("flowFrame", exprs = as.matrix(data_comp[, c])) # 1:22 channels, w/o the labels
ff_uncomp <- new("flowFrame", exprs = as.matrix(data_uncomp[, c]))
ff_comp_eq <- new("flowFrame", exprs = as.matrix(data_comp_eq[, c])) 
ff_uncomp_eq <- new("flowFrame", exprs = as.matrix(data_uncomp_eq[, c]))
ff_comp_diff <- new("flowFrame", exprs = as.matrix(data_comp_diff[, c]))  
ff_uncomp_diff <- new("flowFrame", exprs = as.matrix(data_uncomp_diff[, c]))
summary(ff_comp)
```

```{r}
write.FCS(ff_comp, paste("ff_comp_11_fluo_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_uncomp, paste("ff_non_comp_11_fluo_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_comp_eq, paste("ff_comp_eq_11_fluo_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_uncomp_eq, paste("ff_non_comp_11_eq_fluo_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_comp_diff, paste("ff_comp_diff_11_fluo_", params$experience, "_", params$beads, ".fcs", sep = ""))
write.FCS(ff_uncomp_diff, paste("ff_non_comp_diff_11_fluo_", params$experience, "_", params$beads, ".fcs", sep = ""))
```

Visulization
```{r}
library(ggcyto)
autoplot(ff_comp, "CD56_APCH7", "CD3_PE")
```

```{r}
autoplot(ff_uncomp, "CD56_APCH7", "CD3_PE")
```



## Saving the labels 
```{r}
saveRDS(data_comp[, "label"], paste("labels",  params$experience, sep = "_"))
saveRDS(data_comp_eq[, "label"], paste("labels_eq",  params$experience,  sep = "_"))
saveRDS(data_comp_diff[, "label"], paste("labels_diff",  params$experience,  sep = "_"))
```




