---
title: "Preparing real data: 11 parameters"
output: html_document
params:
  experience: "real_03_003"
  file: "real_03_003_v2.wsp"
  sample: 1
  cells: 100000
  condition: "down_100"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = F)
```

```{r}
library(flowWorkspace)
library(caret)
```

# Introduction
An `R Markdown` script has been developed in order to prepare flow cytometry data for algorithm benchmarking. The data has been previously pre-treated and gated in `FlowJo`. This script enables importing and exploring the data, applying the Boolean strategy performed on `FlowJo`, labelling the examples according to this gating, performing down-sampling and exporting the data for further analysis. 

# Data
This experimental design includes 11 parameters or markers.

# Parameters
```{r}
params
```

# Importing data from FlowJo workspace  (`.wsp`)
The `FlowJo` workspace is open using `openWorkspace()` from the `flowWorkspace` package.
```{r}
(ws <-  openWorkspace(params$file))
```

# Parsing the workspace: GATING SET
```{r}
gs <-  parseWorkspace(ws, name = 2, sampNloc="sampleNode") # 1: All Samples, 2: group
```

```{r}
gs@data
```

Gating strategy:
```{r}
plot(gs)
```

```{r}
png(paste("exports/gating_", params$experience, ".png", sep = ""), res=150, width = 1400, height = 1000)
plot(gs)
dev.off()
```

# Extracting the data from one sample
The sample is defined in the parameters:
```{r}
(s <- params$sample)
```

```{r}
getSamples(ws)[s,]
```

```{r}
getSamples(ws)[s,"name"]
```

Data:
```{r}
gs@data[[s]]
```

```{r}
summary(gs@data[[s]])
```

Counts:
```{r}
getPopStats(gs[[s]])
```

Variation between `FlowJo` and `openCyto` counts:
```{r, fig.height=10}
plotPopCV(gs[[s]])
```

Extracting the data
```{r}
data <- exprs(gs@data[[s]])
# Or, equivalently:
# data  <- getData(gs[[s]])
nrow(data)
```

## Extracting cell labels
Preparing the data frame
```{r}
colnames(data)
```

We are interested in columns 7-12, 14-18:
```{r}
colnames(data)[c(c(7:12), c(14:18))]
```

```{r}
data <- data[,c(c(7:12), c(14:18))]
data <- as.data.frame(data)
```

Getting the nodes (or the populations):
```{r}
(nodes <- getNodes(gs, path = "auto"))
```

Let's see which cell populations we want to label:
```{r}
#getNodes(gs)
```

## CELLS
We are only interested in viable, single cells lymphocytes (no monocytes): node 6.
```{r}
nodes[6]
cells <- getIndices(gs[[s]], nodes[6])
length(cells)
table(cells)
```

Let's tag them as "lymphocytes". The rest, will be "other_events".
```{r}
data$cells <-  "other_events" 
data$cells <- as.factor(ifelse(cells, "lymphocytes", data$cells))
data$cells <- ifelse(cells, "lymphocytes", data$cells)
table(data$cells)
```

## LABELS
19 populations
```{r}
label_list <- c("8" = "NO_BTNK", "16" = "NK", "17" = "NKT", 
                "10" = "B_S_M_27pos", "13" = "B_UNSWCH", "14" = "B_NAIVE", "15" = "B_S_M_27neg",
                "22" = "T_4_EMRA", "23" = "T_4_TN", "27" = "T_4_TCM", "28" = "T_4_TEM", "24" = "T4_RA_nTreg", "29" = "T4_RO_nTreg",
                "31" = "T_8_EMRA", "34" = "T_8_TN", "32" = "T_8_TCM", "33" = "T_8_TEM", 
                "35" = "T_DP", "36" = "T_DN")
names(label_list)
```

Labelling:
```{r}
data$label_list <- "outliers"

for(i in names(label_list)){
    indices <- getIndices(gs[[s]], nodes[as.numeric(i)])
    data$label_list <- ifelse(indices, label_list[[i]], data$label_list)
}

data$label_list <- factor(data$label_list, levels = c(label_list, "outliers"))
table(data$label_list)
```

## LYMPHOCYTES
We are going to keep only the living, single cell lymphocytes:
```{r}
data <- data[data$cells == "lymphocytes",]
nrow(data)
```

# Export data 
## Conversion to `FlowFrame` class objects
```{r}
colnames(data)
```

Replacing the fluorophores by the markers:
```{r}
colnames(data)[1:11] <-  c("CD56", "CD8a_IgD", "CD38", "CD19", "CD127", "CCR7",
                           "CD25", "CD4", "CD3", "CD27", "CD45RA")
colnames(data)
```

## Frequencies
```{r export frequencies}
pops_freq <- (prop.table(table(data$label_list)))*100

saveRDS(pops_freq, paste("pops_freq", params$experience, s, sep = "_")) 
```

## Down-sampling
We use the `createDataPartition()` function from the `caret` package. This function generates partitions with representative frequencies for the indicated classes. We are going to base the partition on the tag "labels" (the subpopulations).
```{r}
p <- params$cells/nrow(data)
subsample <- createDataPartition(data$label_list, p = p, list = F)
data <- data[subsample, ]
nrow(data)
```

## Conversion to `FlowFrame` class objects
```{r}
library(flowCore)
ff <- new("flowFrame", exprs = as.matrix(data[, 1:11])) # 1:11 markers, w/o the labels
summary(ff)
```

Visulaization
```{r}
library(ggcyto)
autoplot(ff[], "CD56", "CD3")
```

## Exporting to `.fcs`
```{r, eval = T}
write.FCS(ff, paste(params$condition, "_ff_", params$experience, ".fcs", sep = ""))
```

## Saving the labels 
```{r}
saveRDS(data[, "label_list"], paste(params$condition, "labels",  params$experience, sep = "_"))
```

## Closing the workspace
```{r}
closeWorkspace(ws)
```


