---
title: "FlowMeans - pop 02"
subtitle: Without previous dimensionality reduction
output:
  html_document:
    number_sections: yes
    theme: journal
    toc: yes
    toc_float: yes
author: Anna Guadall
params:
  file: "ff_02_2019-04-01.fcs"
  pheno: "pheno_02_2019-04-01"
  labels: "labels_02_2019-04-01"
  seed: 23
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(cytofkit)
```

# Reading the data
```{r}
ff <- flowCore::read.FCS(params$file)

# Import the labels
labels <- readRDS(params$labels)

# number of populations
c <- length(levels(labels))
```

Our data matrix:
```{r}
dim(ff@exprs)
```

```{r}
# t-SNE to visualize results
tsne_red <-  cytof_dimReduction(data = ff@exprs, method = "tsne", tsneSeed = params$seed)

data <- cbind(as.data.frame(tsne_red), as.data.frame(labels))
cytof_clusterPlot(data = data, xlab = "tsne_1", ylab = "tsne_2", 
                  cluster = "labels", sampleLabel = FALSE, labelSize = 4)
```

# Partitioning: FlowMeans
k-means + merging of clusters
http://127.0.0.1:28454/library/flowMeans/doc/flowMeans.pdf

### User-defined number of clusters
```{r}
library(flowMeans)
fm <- flowMeans(ff@exprs, MaxN = c )
#  MaxN: maximum number of clusters
# c is the number of cell types in the synthetic sample

table(fm@Labels[[1]])
```

```{r}
data <- as.data.frame(cbind(data, flowmeans = fm@Labels[[1]]))

cytof_clusterPlot(data, xlab = "tsne_1", ylab = "tsne_2", 
                  cluster = "flowmeans", sampleLabel = FALSE, labelSize = 4)
```

#### Evaluation
Matching the labels with the predictions:
```{r}
(t <- table(flow_means = fm@Labels[[1]], labels))
```

```{r}
# sorting max values
max_values <- apply(t, 2, max)
sorted_max_values <- sort(max_values)

# Finding the maximum number of each cell type (columns) on each cluster (rows):
m <- apply(t, 2, which.max)

# Adding the label "unknown" to the labels:
levels(data$labels) <- c(levels(data$labels), "unknown")

data$flowmeans <- "unknown" # all unknowns, unless:

# Replacing the numbers of the clusters by the names of the cell types:
for(i in 1:length(fm@Labels[[1]])){
  for(j in 1:c){ 
    if(fm@Labels[[1]][i] == m[names(sorted_max_values)[j]]){ # we compare to the sorted names
      data$flowmeans[i] <- names(sorted_max_values)[j]
    }
  }
}

data$flowmeans <- factor(data$flowmeans, levels = levels(data$labels))

table(data$flowmeans, data$labels)
```

Here, all B cells are labelled as U4, but it would have been the other way around, all the U4 labelled as B, if the label B had come after U4.  
As we name the clusters following increasing number of cells, the method will favor the matching to the more abundant populations. 

Computing other performance measurements:
```{r}
library(caret)
(cm <- confusionMatrix(data = data$flowmeans, reference = data$labels))

cm$byClass
```

### Number of clusters determined automatically
```{r}
fm_auto <- flowMeans(ff@exprs, MaxN = NA )
#  MaxN: maximum number of clusters

table(fm_auto@Labels[[1]])
```

```{r}
data <- as.data.frame(cbind(data, flowmeans_auto = fm_auto@Labels[[1]]))

cytof_clusterPlot(data,  xlab = "tsne_1", ylab = "tsne_2", 
                  cluster = "flowmeans_auto", sampleLabel = FALSE, labelSize = 4)
```

#### Evaluation
Matching the labels with the predictions:
```{r}
(t <- table(flow_means = fm_auto@Labels[[1]], labels))
```

```{r}
# sorting max values
max_values <- apply(t, 2, max)
sorted_max_values <- sort(max_values)

# Finding the maximum number of each cell type (columns) on each cluster (rows):
m <- apply(t, 2, which.max)

# Adding the label "unknown" to the labels:
#levels(data$labels) <- c(levels(data$labels), "unknown")

data$flowmeans_auto <- "unknown" # all unknowns, unless:

# Replacing the numbers of the clusters by the names of the cell types:
for(i in 1:length(fm_auto@Labels[[1]])){
  for(j in 1:c){ 
    if(fm_auto@Labels[[1]][i] == m[names(sorted_max_values)[j]]){ # we compare to the sorted names
      data$flowmeans_auto[i] <- names(sorted_max_values)[j]
    }
  }
}

data$flowmeans_auto <- factor(data$flowmeans_auto, levels = levels(data$labels))

table(data$flowmeans_auto, data$labels)
```

Computing other performance measurements:
```{r}
library(caret)
(cm <- confusionMatrix(data = data$flowmeans_auto, reference = data$labels))

cm$byClass
```

