---
title: "RPhenograph - test_05"
output:
  html_document:
    number_sections: yes
    theme: journal
    toc: yes
    toc_float: yes
author: Anna Guadall
params:
  file: "ff_05_2019-04-18.fcs"
  labels: "labels_05_2019-04-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

Rphenograph is included in the package cytofkit. It can also be downloaded from the JinmiaoChenLab at GitHub.

# Reading the data
```{r}
ff <- flowCore::read.FCS(params$file)

# Import the labels
labels <- readRDS(params$labels)

# number of populations
c <- length(levels(labels))
```

```{r}
# expression matrix is in ff@exprs
dim(ff@exprs)
```

```{r}
is.matrix(ff@exprs)
```

# RphenoGraph partitioning
```{r}
library(cytofkit)
rphe <- Rphenograph(ff@exprs) # default k = 30 nearest neighbours
```

```{r}
#print(rphe)
```

```{r}
length(rphe)
```

```{r}
# returns a numeric vector, one number for each vertex in the graph that was the input of the community detection
#rphe$membership
```

```{r}
# returns a numeric scalar
rphe$modularity
# ????
```

```{r}
# returns a character scalar
rphe$algorithm
```

```{r}
table(rphe$membership)
```


# Evaluating the performance (NEW MATCHING PROCEDURE)
Matching the labels with the predictions:

```{r, message=FALSE}
# Cross table (CLUSTERING)
(t <- table(rphe$membership, labels))
```

Each cluster is assigned to the label for which it contains a higher number of cells. Several clusters can be assigned to the same label:
```{r, echo = T}
# Finding and sorting row maximums:
max_values <- apply(t, 1, max)
(sorted_max_values <- sort(max_values))
```
\  

```{r, echo = T}
# Finding the cell population (columns) 
# with a higher number of cells for each cluster (rows):
(m <- apply(t, 1, which.max)) 
```
\  

```{r, echo = T}
table(m)
```

Matching cell labels and clusters:
```{r, echo = T}
# Empty list 
matched_preds <- rep("NA", length(rphe$membership))

# Number of predicted clusters
nc <- length(m)

# Replacing the numbers of the clusters by the names of the cell types:
for(i in 1:length(rphe$membership)){
  for(j in 1:nc){
    if(rphe$membership[i] == names(sorted_max_values)[j]){
      matched_preds[i] <- levels(labels)[m[[names(sorted_max_values)[j]]]] 
        # m : list of the more represented cell labels for every cluster
        # m[[names(sorted_max_values)[j]]] : 
            # Label index assigned for that cluster
        # levels(labels)[m[[names(sorted_max_values)[j]]]] : 
            # Label name assigned for that index
    }
  }
}
  
# Factorize matched predictions
matched_preds <- factor(matched_preds, levels = levels(labels))
table(matched_preds, labels)
```

Computing the confusion matrix and other performance measurements:
```{r}
library(caret)
(cm <- confusionMatrix(data = matched_preds, reference = labels))
```

```{r}
cm$byClass
```

# Which populations have been split?
```{r, echo = T}
(table_maxs <- table(m))
```

```{r}
table_maxs[table_maxs>1]
```

```{r}
(split_list <-  names(table_maxs[table_maxs>1]))
```

Which population is that?
```{r}
# split_list is character
split_list <-  as.numeric(split_list)

levels(labels)[split_list] 
```

How many clusters have been predicted for this population?
```{r}
table_maxs[[split_list]]
```

# Visualization

```{r}
dades <- cbind(as.data.frame(ff@exprs), labels, matched_preds)
ggplot(dades, aes(x = CD3, y = CD19, col = labels, shape = matched_preds))  + 
  geom_point(size = 3) + scale_shape_manual(values = c(0:10)) +
  theme(legend.position = "bottom")
```

```{r}
ggplot(dades, aes(x = CD8, y = CD56, col = labels, shape = matched_preds))  + 
  geom_point(size = 3) + scale_shape_manual(values = c(0:10)) +
  theme(legend.position = "bottom")
```





