---
title: "RPhenograph - pop 02"
output:
  html_document:
    number_sections: yes
    theme: journal
    toc: yes
    toc_float: yes
author: Anna Guadall
params:
  file: "ff_02_2019-04-01.fcs"
  pheno: "pheno_02_2019-04-01"
  labels: "labels_02_2019-04-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Rphenograph is included in the package cytofkit. It can also be downloaded from the JinmiaoChenLab at GitHub.

# Reading the data
```{r}
ff <- flowCore::read.FCS(params$file)

# Import the labels
labels <- readRDS(params$labels)

# number of populations
c <- length(levels(labels))
```

```{r}
# expression matrix is in ff@exprs
dim(ff@exprs)
```

```{r}
is.matrix(ff@exprs)
```

# RphenoGraph partitioning
```{r}
library(cytofkit)
rphe <- Rphenograph(ff@exprs) # default k = 30 nearest neighbours
```

```{r}
print(rphe)
```

```{r}
length(rphe)
```

The algorithm has found 11 groups. Great!

```{r}
# returns a numeric vector, one number for each vertex in the graph that was the input of the community detection
rphe$membership
```

Group numbers have been given consecutively.

```{r}
# returns a numeric scalar
rphe$modularity
# ????
```

```{r}
# returns a character scalar
rphe$algorithm
```

```{r}
table(rphe$membership)
```

Just perfect. But let's verify it anyway.  

# Evaluating the performance

Matching the labels with the predictions:

```{r}
#cells <- as.data.frame(cbind(cells = seq(1, length(labels), 1), mc_cell, labels))
cells <- cbind(as.data.frame(labels), pred = rphe$membership)
head(cells)
```

```{r, message=FALSE}
(t <- table(cells$pred, cells$labels))
```

Finding the maximum number of each cell type (columns) on each cluster (rows):
```{r}
(m <- apply(t, 2, which.max))
```

Replacing the numbers of the clusters by the names of the cell types:
```{r}
# on the cell metaclustering
for(i in 1:length(cells$pred)){
  for(j in 1:c){
    if(cells$pred[i] == m[[j]]){
      cells$pred[i] <- levels(labels)[j] 
    }
  }
}

table(cells$pred, cells$labels)
```

This is the confusion matrix.

Computing the confusion matrix and other performance measurements:
```{r}
library(caret)
cells$pred <- factor(cells$pred, levels = levels(labels))
(cm <- confusionMatrix(data = cells$pred, reference = labels))
```

```{r}
cm$byClass
```


```{r}
dades <- cbind(as.data.frame(ff@exprs), cells$labels, cells$pred)
ggplot(dades, aes(x = CD3, y = CD19, col = labels, shape = cells$pred))  + 
  geom_point(size = 3) + scale_shape_manual(values = c(0:10)) +
  theme(legend.position = "bottom")
```

```{r}
ggplot(dades, aes(x = CD8, y = CD56, col = labels, shape = cells$pred))  + 
  geom_point(size = 3) + scale_shape_manual(values = c(0:10)) +
  theme(legend.position = "bottom")
```





