---
title: "test flowSOM - synthetic sample"
output:
  html_document:
    number_sections: yes
    theme: journal
    toc: yes
    toc_float: yes
author: Anna Guadall
params:
  file: "ff_02_2019-04-01.fcs"
  pheno: "pheno_02_2019-04-01"
  labels: "labels_02_2019-04-01"
  seed: 23
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reading the data
```{r}
# import the fcs file
#ff <- flowCore::read.FCS(params$file)

# Import the labels
labels <- readRDS(params$labels)

# number of populations
c <- length(levels(labels))
```

# The easy way (reading and clustering)
Creating a large list from an FCS file with the wrapper `FlowSOM`  
No scaling:
```{r}
#library(flowCore)
#fframe <- read.FCS(params$file, transformation = F)

library(FlowSOM)
fs <- FlowSOM(params$file,
                # Input options:
                compensate = F,transform = F,
                scale = F,
                # SOM options:
                colsToUse = colnames(params$file), 
                #xdim = 7, ydim = 7, # determines the number of nodes and dimensions of the grid
                # Metaclustering options:
                nClus = 11,
                # Seed for reproducible results:
                seed = params$seed)

PlotStars(fs$FlowSOM, backgroundValues = as.factor(fs$metaclustering))
```

```{r}
PlotStars(fs$FlowSOM, backgroundValues = as.factor(fs$metaclustering), view = "grid")
```

# Step by step
## Reading the data
```{r}
ff <- suppressWarnings(flowCore::read.FCS(params$file))
ff
```

```{r}
fs <- ReadInput(ff, compensate = F, transform = F, scale = F)
# compensate = F, transform = F : there is no compensation matrix

str(fs, max.level = 2)
```

## Building the self-organizing map

```{r, eval = FALSE}
# Which columns?
names(fs$prettyColnames)
```

In this case, these are all the columns, no need to specify.
```{r}
set.seed(params$seed) # set the seed! The same used in "the easy way" if we want to have the same results
fs <- BuildSOM(fs)
str(fs$map,max.level = 2)
```

## Building the minimal spanning tree
**BuildMST()** will return a FLowSOM object with extra information contained in the $MST parameter.
```{r}
fs <- BuildMST(fs,tSNE=TRUE)
str(fs$MST)
```

## Plotting
### Minimal Spanning Tree
```{r}
PlotStars(fs)
```

### SOM grid
```{r}
PlotStars(fs, view="grid")
```

### tSNE
Only possible when tSNE was TRUE in BuildMST.
```{r}
PlotStars(fs,view="tSNE")
```

## Looking just at one specific marker with `PlotMarker`
```{r}
print(colnames(fs$map$medianValues))
```

```{r}
for(i in 1:length(colnames(fs$map$medianValues))){
  PlotMarker(fs, colnames(fs$map$medianValues)[[i]])
}
```

```{r}
for(i in 1:length(colnames(fs$map$medianValues))){
  PlotMarker(fs, colnames(fs$map$medianValues)[[i]], view = "tSNE")
}
```

## Numbering the nodes
```{r}
PlotNumbers(UpdateNodeSize(fs,reset=TRUE), nodeSize = 15)
```

```{r}
PlotNumbers(UpdateNodeSize(fs,reset=TRUE), nodeSize = 20, view = "grid")
```

```{r}
PlotNumbers(UpdateNodeSize(fs,reset=TRUE), nodeSize = 15, view = "tSNE")
```

```{r}
PlotClusters2D(fs, "CD3","CD19",c(8, 9, 10, 18, 19, 20))
```



## Meta-clustering the data
This can be the first step in further analysis of the data, and often gives a good approximation of manual gating results.

```{r}
head(fs$map$codes)
dim(fs$map$codes)
```

100 nodes, 5 markers.

K: Number of clusters
```{r}
mc <- metaClustering_consensus(fs$map$codes, k = c) # c: number of cell types on the synthetic data
mc
```

```{r}
PlotStars(fs, backgroundValues = as.factor(mc))
```

```{r}
PlotStars(fs,view="grid", backgroundValues = as.factor(mc))
```

```{r}
PlotStars(fs,view="tSNE", backgroundValues = as.factor(mc))
```

### Meta-clustering each cell individually
```{r}
head(fs$map$mapping)
```

```{r}
dim(fs$map$mapping)
```
There are `r nrow(fs$map$mapping)` files, one per cell.  

```{r}
summary(fs$map$mapping[,1])
# table(fs$map$mapping[,1])
```

Column 1 assigns one node to each cell (`r max(fs$map$mapping[,1])` nodes).

```{r}
summary(fs$map$mapping[,2])
```

I don't know what is on column 2.

```{r}
mc_cell <- mc[fs$map$mapping[,1]]
mc_cell[1:10]
```

```{r}
table(mc_cell)
```

It looks nice. As I now how my "cells" are ordered, I can verify for every type of them:

B cells:
```{r}
n <- 909
table(mc_cell[1:n])
```

NK cells:
```{r}
table(mc_cell[(n+1):(n+n)])
```

T4 cells:
```{r}
t <- n + n
table(mc_cell[(t+1):(t+n)])
```

T8 cells:
```{r}
t <- t + n
table(mc_cell[(t+1):(t+n)])
```

NKT_NN cells:
```{r}
t <- t + n
table(mc_cell[(t+1):(t+n)])
```

NKT_4 cells:
```{r}
t <- t + n
table(mc_cell[(t+1):(t+n)])
```

NKT_8 cells:
```{r}
t <- t + n
table(mc_cell[(t+1):(t+n)])
```

U1 cells:
```{r}
t <- t + n
table(mc_cell[(t+1):(t+n)])
```

U2 cells:
```{r}
t <- t + n
table(mc_cell[(t+1):(t+n)])
```

U3 cells:
```{r}
t <- t + n
table(mc_cell[(t+1):(t+n)])
```

U4 cells:
```{r}
t <- t + n
table(mc_cell[(t+1):(t+n)])
```

#### Evaluating the performance

Matching the labels with the predictions:

```{r}
#cells <- as.data.frame(cbind(cells = seq(1, length(labels), 1), mc_cell, labels))
cells <- cbind(as.data.frame(labels), mc_cell)
head(cells)
```

```{r, message=FALSE}
(t <- table(cells$mc_cell, cells$labels))
```

Finding the maximum number of each cell type (columns) on each cluster (rows):
```{r}
(m <- apply(t, 2, which.max))
```

Replacing the numbers of the clusters by the names of the cell types:
```{r}
# on the cell metaclustering
for(i in 1:length(mc_cell)){
  for(j in 1:c){
    if(cells$mc_cell[i] == m[[j]]){
      cells$mc_cell[i] <- levels(labels)[j] 
    }
  }
}

table(cells$mc_cell, cells$labels)
```

This is the confusion matrix.

Replacing the numbers of the clusters by the names of the cell types:
```{r}
table(mc)
```

```{r}
# on the node metaclustering
for(i in 1:length(mc)){
  for(j in 1:c){
    if(mc[i] == m[[j]]){
      mc[i] <- levels(labels)[j] 
    }
  }
}

table(mc)
```

Computing the confusion matrix and other performance measurements:
```{r}
library(caret)
cells$mc_cell <- factor(cells$mc_cell, levels = levels(labels))
(cm <- confusionMatrix(data = cells$mc_cell, reference = labels))
```

```{r}
cm$byClass
```

Great! 

We can visualize the original cell types on the MST:  
Original cell types are represented as plot pies.  
We can add the result of the metaclustering on the background.
```{r}
PlotPies(fs, cellTypes = labels, backgroundValues = as.factor(mc))
```

Or on the star plots:
```{r}
PlotStars(fs, backgroundValues = as.factor(mc))
```

Or on the heat maps:
```{r}
for(i in 1:length(colnames(fs$map$medianValues))){
  PlotMarker(fs, colnames(fs$map$medianValues)[[i]], backgroundValues = as.factor(mc))
}
```

```{r}
for(i in 1:length(colnames(fs$map$medianValues))){
  PlotMarker(fs, colnames(fs$map$medianValues)[[i]], view = "tSNE", backgroundValues = as.factor(mc))
}
```

So many colors...

## Detecting nodes with a specific pattern
I think this functionality will not be very useful.
```{r}
# Import the patterns
patterns <- readRDS(params$pheno)
patterns
```

```{r}
# All nodes as "unknown"
cellTypes_01 <- factor(rep("Unknown",100),levels=c("Unknown", rownames(patterns)))

for(i in rownames(patterns)){
  query <-  QueryStarPlot(UpdateNodeSize(fs, reset=T), patterns[i,], plot = F)
  cellTypes_01[query$selected] <- i
}

PlotStars(fs, backgroundValues=cellTypes_01)
```


```{r}
PlotStars(fs,view="tSNE", backgroundValues = cellTypes_01)
```






