---
title: "t-SNE - pop 02"
output:
  html_document:
    number_sections: yes
    theme: journal
    toc: yes
    toc_float: yes
author: Anna Guadall
params:
  file: "ff_02_2019-04-01.fcs"
  pheno: "pheno_02_2019-04-01"
  labels: "labels_02_2019-04-01"
  seed: 23
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = F)
```

# Reading the data
```{r}
ff <- flowCore::read.FCS(params$file)

# Import the labels
labels <- readRDS(params$labels)

# number of populations
c <- length(levels(labels))
```

# t-SNE: Dimensionality reduction
Our data matrix:
```{r}
dim(ff@exprs)
```

We use t-SNE from the **Cytofkit** package:  

https://www.bioconductor.org/packages/3.7/bioc/vignettes/cytofkit/inst/doc/cytofkit_example.html
```{r}
library(cytofkit)
# in the example, data is transformed prior to t-SNE reduction
tsne_red <-  cytof_dimReduction(data = ff@exprs, method = "tsne", tsneSeed = params$seed)
# default seed : 42 

head(tsne_red)
```

```{r}
plot(tsne_red)
```

```{r}
# combine data
data <- cbind(as.data.frame(tsne_red), as.data.frame(labels))
head(data)
```

```{r}
cytof_clusterPlot(data = data, xlab = "tsne_1", ylab = "tsne_2", 
                  cluster = "labels", sampleLabel = FALSE, labelSize = 4)
```

# Partitioning
## ClusterX: Density peak-finding
```{r}
cluster_x <-  cytof_cluster(tsne_red, method = "ClusterX")
# the number of clusters cannot be fixed
```

```{r}
table(cluster_x)
```

```{r}
# combine data
data <- as.data.frame(cbind(data, ClusterX = cluster_x))
head(data)
```

```{r}
cytof_clusterPlot(data=data, xlab="tsne_1", ylab="tsne_2",
                  cluster="ClusterX", sampleLabel = F, labelSize = 4)
```

Dots can be replaed by sample labels:
```{r}
cytof_clusterPlot(data = data, xlab="tsne_1", ylab="tsne_2",  sample = "labels",
                  cluster="ClusterX", sampleLabel = T, labelSize = 4)
```

### Evaluation
Matching the labels with the predictions:
```{r}
(t <- table(cluster_x, labels))
```

Finding the maximum number of each cell type (columns) on each cluster (rows):
```{r}
(m <- apply(t, 2, which.max))
```

Replacing the numbers of the clusters by the names of the cell types:
```{r}
for(i in 1:length(cluster_x)){
  for(j in 1:c){
    if(cluster_x[i] == m[[j]]){
      data$ClusterX[i] <- levels(labels)[j] 
    }
  }
}

table(data$ClusterX, data$labels)
```

What can we do with the extra labels? Maybe I could class them as "unknown". I should also add this class to the labels list:
```{r}
# Adding the label "unknown" to the labels:
levels(data$labels) <- c(levels(data$labels), "unknown")

(t <- table(cluster_x, data$labels))
```

```{r}
m <- apply(t, 2, which.max)

data$ClusterX <-  "unknown" # all unknowns, unless:

for(i in 1:length(cluster_x)){
  for(j in 1:c){
    if(cluster_x[i] == m[[j]]){
      data$ClusterX[i] <- levels(labels)[j] # labels has 11 levels
    }
  }
}

table(data$ClusterX, data$labels) # data$labels has 12 levels
```

This is the confusion matrix.  
Computing other performance measurements:
```{r}
data$ClusterX <- factor(data$ClusterX, levels = levels(data$labels))

library(caret)
(cm <- confusionMatrix(data = data$ClusterX, reference = data$labels))

cm$byClass
```

The F1 scores can be found on:
```{r, eval=FALSE}
cm$byClass[,7]
```

## FlowMeans
k-means + merging of clusters

### User-defined number of clusters
```{r}
library(flowMeans)
fm <- flowMeans(tsne_red, MaxN = c )
#  MaxN: maximum number of clusters
# c is the number of cell types in the synthetic sample

table(fm@Labels[[1]])
```

```{r}
data <- as.data.frame(cbind(data, flowmeans = fm@Labels[[1]]))

cytof_clusterPlot(data, xlab="tsne_1", ylab="tsne_2", 
                  cluster = "flowmeans", sampleLabel = FALSE, labelSize = 4)
```

#### Evaluation
Matching the labels with the predictions:
```{r}
(t <- table(flowmeans = fm@Labels[[1]], labels))
```

```{r}
m <- apply(t, 2, which.max)

data$flowmeans <-  "unknown" # all unknowns, unless:

for(i in 1:length(fm@Labels[[1]])){
  for(j in 1:c){
    if(fm@Labels[[1]][i] == m[[j]]){
      data$flowmeans[i] <- levels(labels)[j] 
    }
  }
}

data$flowmeans <- factor(data$flowmeans, levels = levels(data$labels))

table(flowmeans = data$flowmeans, labels = data$labels)
```

This is the confusion matrix.  
Computing other performance measurements:
```{r}
library(caret)
(cm <- confusionMatrix(data = data$flowmeans, reference = data$labels))

cm$byClass
```

### Number of clusters determined automatically
```{r}
fm_auto <- flowMeans(tsne_red, MaxN = NA )
#  MaxN: maximum number of clusters

table(fm_auto@Labels[[1]])
```

```{r}
data <- as.data.frame(cbind(data, flowmeans_auto = fm_auto@Labels[[1]]))

cytof_clusterPlot(data, xlab="tsne_1", ylab="tsne_2", 
                  cluster = "flowmeans_auto", sampleLabel = FALSE, labelSize = 4)
```

#### Evaluation
Matching the labels with the predictions:
```{r}
(t <- table(flowmeans = fm_auto@Labels[[1]], labels))
```

```{r}
m <- apply(t, 2, which.max)

data$flowmeans_auto <- "unknown" # all unknowns, unless:

for(i in 1:length(fm_auto@Labels[[1]])){
  for(j in 1:c){
    if(fm_auto@Labels[[1]][i] == m[[j]]){
      data$flowmeans_auto[i] <- levels(labels)[j] 
    }
  }
}

data$flowmeans_auto<- factor(data$flowmeans_auto, levels = levels(data$labels))

table(data$flowmeans_auto, data$labels)
```

There is a problem here. B max is on cluster 9. But THEN, T4 max is also in cluster 9. So the B cells are matched as T4. How could we solve that? The maximum values should be oredered increasingly:
```{r}
(max_values <- apply(t, 2, max))
```

```{r}
m
```

```{r}
(sorted_max_values <- sort(max_values))
```

```{r}
m <- apply(t, 2, which.max)

data$flowmeans_auto <- "unknown" # all unknowns, unless:

for(i in 1:length(fm_auto@Labels[[1]])){
  for(j in 1:c){ 
    n <- names(sorted_max_values)[j] # we compare to the sorted names
    if(fm_auto@Labels[[1]][i] == m[n]){ 
      data$flowmeans_auto[i] <- n
    }
  }
}

data$flowmeans_auto <- factor(data$flowmeans_auto, levels = levels(data$labels))

table(data$flowmeans_auto, data$labels)
```

We should remember the plots:
```{r}
cytof_clusterPlot(data, xlab="tsne_1", ylab="tsne_2", 
                  cluster = "labels", sampleLabel = FALSE, labelSize = 4)
cytof_clusterPlot(data, xlab="tsne_1", ylab="tsne_2", 
                  cluster = "flowmeans_auto", sampleLabel = FALSE, labelSize = 4)
```

Now the match is correct.

Computing other performance measurements:
```{r}
library(caret)
(cm <- confusionMatrix(data = data$flowmeans_auto, reference = data$labels))

cm$byClass
```


## FlowPeaks
k-means + finite mixture model

```{r}
library(flowPeaks)
red_data <- as.matrix(tsne_red)
colnames(red_data) <- c("tsne_1", "tsne_2")
fp <-  flowPeaks(red_data)
```

```{r}
table(fp$peaks.cluster)
```

```{r}
plot(fp)
```

```{r}
data <- as.data.frame(cbind(data, flow_peaks = fp$peaks.cluster))

cytof_clusterPlot(data, xlab="tsne_1", ylab="tsne_2", 
                  cluster = "flow_peaks", sampleLabel = FALSE, labelSize = 4)
```

### Evaluation
Matching the labels with the predictions:
```{r}
(t <- table(flow_peaks = fp$peaks.cluster, labels))
```

```{r}
(max_values <- apply(t, 2, max))
```

```{r}
(sorted_max_values <- sort(max_values))
```

```{r}
m <- apply(t, 2, which.max)

data$flow_peaks <- "unknown" # all unknowns, unless:

for(i in 1:length(fp$peaks.cluster)){
  for(j in 1:c){ 
    if(fp$peaks.cluster[i] == m[names(sorted_max_values)[j]]){ # we compare to the sorted names
      data$flow_peaks[i] <- names(sorted_max_values)[j]
    }
  }
}

data$flow_peaks <- factor(data$flow_peaks, levels = levels(data$labels))

table(data$flow_peaks, data$labels)
```

Computing other performance measurements:
```{r}
library(caret)
(cm <- confusionMatrix(data = data$flow_peaks, reference = data$labels))

cm$byClass
```



