---
title: "UMAP - pop 02"
output:
  html_document:
    number_sections: yes
    theme: journal
    toc: yes
    toc_float: yes
author: Anna Guadall
params:
  file: "ff_02_2019-04-01.fcs"
  pheno: "pheno_02_2019-04-01"
  labels: "labels_02_2019-04-01"
  seed: 23
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(cytofkit)
```

# Reading the data
```{r}
ff <- flowCore::read.FCS(params$file)

# Import the labels
labels <- readRDS(params$labels)

# number of populations
c <- length(levels(labels))
```

# UMAP: Dimensionality reduction
https://cran.r-project.org/web/packages/umap/vignettes/umap.html  

Our data matrix:
```{r}
dim(ff@exprs)
```

```{r}
library(umap)
(umap_red <- umap(ff@exprs, random_state = params$seed))
# default seed: 42
```

```{r}
head(umap_red$layout)
```

```{r}
plot(umap_red$layout)
```

```{r}
umap.defaults
```

```{r}
# combine data
data <- cbind(as.data.frame(umap_red$layout), as.data.frame(labels))
head(data)
```

```{r}
cytof_clusterPlot(data = data, xlab = "V1", ylab = "V2", 
                  cluster = "labels", sampleLabel = F, labelSize = 4)
```

# Partitioning
## ClusterX: Density peak-finding
```{r}
library(cytofkit)
cluster_x <-  cytof_cluster(umap_red$layout, method = "ClusterX")
# the number of clusters cannot be fixed
```

```{r}
table(cluster_x)
```

```{r}
# combine data
data <- as.data.frame(cbind(data, ClusterX = cluster_x))
head(data)
```

```{r}
cytof_clusterPlot(data = data, xlab = "V1", ylab = "V2", 
                  cluster = "ClusterX", sampleLabel = F, labelSize = 4)
```

ClusterX works perfectly with seed = 23. But I had first tested it with with seed = 42 and it was terrible...

```{r}
library(umap)
umap_42 <- umap(ff@exprs, random_state = 42)
library(cytofkit)
cluster_x_42 <-  cytof_cluster(umap_42$layout, method = "ClusterX")
data_42 <- as.data.frame(cbind(umap_42$layout, ClusterX_42 = cluster_x_42))
cytof_clusterPlot(data = data_42, xlab = "V1", ylab = "V2", 
                  cluster = "ClusterX_42", sampleLabel = F, labelSize = 4)
```

### Evaluation
Matching the labels with the predictions:
```{r}
(t <- table(cluster_x, labels))
```

```{r}
# sorting max values
max_values <- apply(t, 2, max)
sorted_max_values <- sort(max_values)

# Finding the maximum number of each cell type (columns) on each cluster (rows):
m <- apply(t, 2, which.max)

# Adding the label "unknown" to the labels:
levels(data$labels) <- c(levels(data$labels), "unknown") # JUST ONCE!
# should do it at the beginning

data$ClusterX <- "unknown" # all unknowns, unless:

# Replacing the numbers of the clusters by the names of the cell types:
for(i in 1:length(cluster_x)){
  for(j in 1:c){ 
    if(cluster_x[i] == m[names(sorted_max_values)[j]]){ # we compare to the sorted names
      data$ClusterX[i] <- names(sorted_max_values)[j]
    }
  }
}

data$ClusterX <- factor(data$ClusterX, levels = levels(data$labels))

table(data$ClusterX, data$labels)
```

Computing other performance measurements:
```{r}
library(caret)
(cm <- confusionMatrix(data = data$ClusterX, reference = data$labels))

cm$byClass
```


## FlowMeans
k-means + merging of clusters

### User-defined number of clusters
```{r}
library(flowMeans)
fm <- flowMeans(umap_red$layout, MaxN = c )
#  MaxN: maximum number of clusters
# c is the number of cell types in the synthetic sample

table(fm@Labels[[1]])
```

```{r}
data <- as.data.frame(cbind(data, flowmeans = fm@Labels[[1]]))

cytof_clusterPlot(data, xlab="V1", ylab="V2", 
                  cluster = "flowmeans", sampleLabel = FALSE, labelSize = 4)
```

#### Evaluation
Matching the labels with the predictions:
```{r}
(t <- table(flow_means = fm@Labels[[1]], labels))
```

```{r}
# sorting max values
max_values <- apply(t, 2, max)
sorted_max_values <- sort(max_values)

# Finding the maximum number of each cell type (columns) on each cluster (rows):
m <- apply(t, 2, which.max)

# Adding the label "unknown" to the labels:
#levels(data$labels) <- c(levels(data$labels), "unknown")

data$flowmeans <- "unknown" # all unknowns, unless:

# Replacing the numbers of the clusters by the names of the cell types:
for(i in 1:length(fm@Labels[[1]])){
  for(j in 1:c){ 
    if(fm@Labels[[1]][i] == m[names(sorted_max_values)[j]]){ # we compare to the sorted names
      data$flowmeans[i] <- names(sorted_max_values)[j]
    }
  }
}

data$flowmeans <- factor(data$flowmeans, levels = levels(data$labels))

table(data$flowmeans, data$labels)
```

Here, all B cells are labelled as U4, but it would have been the other way around, all the U4 labelled as B, if the label B had come after U4.  
As we name the clusters following increasing number of cells, the method will favor the matching to the more abundant populations. 

Computing other performance measurements:
```{r}
library(caret)
(cm <- confusionMatrix(data = data$flowmeans, reference = data$labels))

cm$byClass
```

### Number of clusters determined automatically
```{r}
fm_auto <- flowMeans(umap_red$layout, MaxN = NA )
#  MaxN: maximum number of clusters

table(fm_auto@Labels[[1]])
```

```{r}
data <- as.data.frame(cbind(data, flowmeans_auto = fm_auto@Labels[[1]]))

cytof_clusterPlot(data, xlab="V1", ylab="V2", 
                  cluster = "flowmeans_auto", sampleLabel = FALSE, labelSize = 4)
```

#### Evaluation
Matching the labels with the predictions:
```{r}
(t <- table(flow_means = fm_auto@Labels[[1]], labels))
```

```{r}
# sorting max values
max_values <- apply(t, 2, max)
sorted_max_values <- sort(max_values)

# Finding the maximum number of each cell type (columns) on each cluster (rows):
m <- apply(t, 2, which.max)

# Adding the label "unknown" to the labels:
#levels(data$labels) <- c(levels(data$labels), "unknown")

data$flowmeans_auto <- "unknown" # all unknowns, unless:

# Replacing the numbers of the clusters by the names of the cell types:
for(i in 1:length(fm_auto@Labels[[1]])){
  for(j in 1:c){ 
    if(fm_auto@Labels[[1]][i] == m[names(sorted_max_values)[j]]){ # we compare to the sorted names
      data$flowmeans_auto[i] <- names(sorted_max_values)[j]
    }
  }
}

data$flowmeans_auto <- factor(data$flowmeans_auto, levels = levels(data$labels))

table(data$flowmeans_auto, data$labels)
```

Computing other performance measurements:
```{r}
library(caret)
(cm <- confusionMatrix(data = data$flowmeans_auto, reference = data$labels))

cm$byClass
```

## FlowPeaks
k-means + finite mixture model

```{r}
library(flowPeaks)
red_data <- as.matrix(umap_red$layout)
colnames(red_data) <- c("V1", "V2")
fp <-  flowPeaks(red_data)

table(fp$peaks.cluster)
```

```{r}
plot(fp)
```

```{r}
data <- as.data.frame(cbind(data, flow_peaks = fp$peaks.cluster))

cytof_clusterPlot(data, xlab="V1", ylab="V2", 
                  cluster = "flow_peaks", sampleLabel = FALSE, labelSize = 4)
```

### Evaluation
Matching the labels with the predictions:
```{r}
(t <- table(flow_peaks = fp$peaks.cluster, labels))
```

```{r}
# sorting max values
max_values <- apply(t, 2, max)
sorted_max_values <- sort(max_values)

# Finding the maximum number of each cell type (columns) on each cluster (rows):
m <- apply(t, 2, which.max)

data$flow_peaks <- "unknown" # all unknowns, unless:

# Replacing the numbers of the clusters by the names of the cell types:
for(i in 1:length(fp$peaks.cluster)){
  for(j in 1:c){ 
    if(fp$peaks.cluster[i] == m[names(sorted_max_values)[j]]){ # we compare to the sorted names
      data$flow_peaks[i] <- names(sorted_max_values)[j]
    }
  }
}

data$flow_peaks <- factor(data$flow_peaks, levels = levels(data$labels))

table(data$flow_peaks, data$labels)
```

Computing other performance measurements:
```{r}
library(caret)
(cm <- confusionMatrix(data = data$flow_peaks, reference = data$labels))

cm$byClass
```




