---
title: "Algorithm benchmarking script"
subtitle: "03_bench_03"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: hide
    theme: united
    highlight: tango
author: Anna Guadall
params:
  threshold: ""
  row: ""
  pg: ""      # phenograph
  fm: ""      # flowMeans
  fp: ""      # flowPeaks
  dp: ""      # DEPECHE
  tpg: ""     # t-SNE + phenograph
  tfm: ""     # t-SNE + flowMeans
  tfp: ""     # t-SNE + flowPeaks
  tdp: ""     # t-SNE + DEPECHE
  tcx: ""     # t-SNE + ClusterX
  upg: ""     # UMAP + phenograph
  ufm: ""     # UMAP + flowMeans
  ufp: ""     # UMAP + flowPeaks
  udp: ""     # UMAP + DEPECHE
  ucx: ""     # UMAP + ClusterX
  out: 1
  norm: ""
  id: ""
  tsne: ""    # t-SNE
  umap: ""    # UMAP
---

```{r setup, include=FALSE, cache = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, cache = T)
```

```{r libraries, message=FALSE, cache = FALSE}
library(caret)
library(cytofkit)
library(flowPeaks)
library(flowMeans)
library(umap)
library(knitr)
library(DepecheR)
```

# Path
```{r path}
getwd()
```

# Parameters
```{r params, cache = F}
params
```

# Import data and functions
```{r import rds objects, cache = F}
ff <- readRDS(list.files(pattern = "ff"))
matching <- readRDS(list.files(pattern = "matching"))
mean_f1 <- readRDS(list.files(pattern = "mean_f1"))
labels <- readRDS(list.files(pattern = "labels"))
summ <- readRDS(list.files(pattern = "summary_table"))
n_pops <- readRDS(list.files(pattern = "n_pops"))
```

**Data: ``r params$id``**    
**`r n_pops` populations.**  
**Normalization: `r params$norm`**

```{r normalize, cache = F}
normalize <- function(x) { return ((x - min(x)) / (max(x) - min(x))) }
```

```{r row, cache = F}
row <- as.numeric(params$row)
```

# Reduction of dimensionality
## t-SNE dimensionality reduction
Perform t-SNE if at least one t-SNE method is required:
```{r decide tsne, cache = F}
do_tsne <- "n"
if (params$tpg == "y" | params$tfm == "y" | params$tfp == "y" | params$tdp == "y" | params$tcx == "y"
    | params$tsne == "y") {
  (do_tsne <- "y") 
} else {
  cat("t-SNE method has not been requested")
}
```

```{r tsne}
if (do_tsne == "y") {
  ptm <- proc.time() # measuring CPU time
  tsne_red <-  cytof_dimReduction(data = ff@exprs, method = "tsne", tsneSeed = 42)
  tsne_time <- proc.time() - ptm
  
  tsne_red <- as.data.frame(tsne_red)
}
```

## UMAP dimensionality reduction
```{r decide umap, cache = F}
do_umap <- "n"
if (params$upg == "y" | params$ufm == "y" | params$ufp == "y" | params$udp == "y" | params$ucx == "y"
    | params$umap == "y") {
  (do_umap <- "y")
} else {
  cat("UMAP method has not been requested")
}
```

```{r umap}
if (do_umap == "y") {
  ptm <- proc.time() # measuring CPU time
  umap_red <- umap(ff@exprs, random_state = 42)
  umap_time <- proc.time() - ptm
  
  colnames(umap_red$layout) <- c("UMAP_1", "UMAP_2")
}
```

## Export dimensionality reduction results
```{r save tsne result, cache = F}
if (do_tsne == "y") {
  saveRDS(tsne_red, "tsne_red.rds")
}
```

```{r save umap result, cache = F}
if (do_umap == "y") {
  saveRDS(umap_red$layout, "umap_red.rds")
}
```

# Clustering on all dimensions
##  RphenoGraph
```{r phenograph clustering}
if (params$pg == "y"){
  method <- "PhenoGraph"
  
  # clustering
  ptm <- proc.time() # measuring CPU time
  pred <- Rphenograph(ff@exprs) # default k = 30 nearest neighbours
  ptm <- proc.time() - ptm
  
  clustering <- pred$membership
} 
```

```{r save clustering, cache = F}
if (params$pg == "y"){
  saveRDS(clustering, "clustering_pheno.rds")
}
```

```{r phenograph matching}
if (params$pg == "y"){
  # MATCHING LABELS AND PREDICTIONS
  matched <- matching(clustering, labels, threshold) 
}
```

```{r save matching, cache = F}
if (params$pg == "y"){
 saveRDS(matched, "matched_pheno.rds") 
}
```

```{r phenograph performance}
if (params$pg == "y"){
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                        reference = labels[labels != "outliers"])
  cm_merged <- confusionMatrix(
    data = matched$merged_clusters[matched$merged_labels != "outliers"],
    reference = matched$merged_labels[matched$merged_labels != "outliers"]
  )
  
  # COMPUTE MEAN F1 and WEIGHTED MEAN F1
  mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops) 
}
```

```{r phenograph results, cache = FALSE}
if (params$pg == "y"){
  summ[row, c(1,2,4)] <- c(params$id, params$norm, method)
  summ[row, -c(1,2,4)] <- c(nrow(ff@exprs), threshold, length(table(clustering)),
                         matched$partitions,
                         mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                         NA, NA, ptm[1] + ptm[4], ptm[3])
  row <- row + 1
  
  # FOR VISUALIZATION
  cat("\n", method)
  cat("\nContingency matrix")
  print(matched$c)
  cat("\nF1 matrix")
  print(round(matched$f, 2))
  cat("\nConfusion matrix")
  print(cm$table)
  cat("\nConfusion matrix (MERGED)")
  print(cm_merged$table) 
}
```

##  flowMeans
```{r flowmeans clustering}
if (params$fm == "y"){
  method <- "flowMeans"
  
  # clustering
  ptm <- proc.time() # measuring CPU time
  pred <- flowMeans(ff@exprs, MaxN = n_pops )
  ptm <- proc.time() - ptm
  
  clustering <- pred@Labels[[1]] 
}
```

```{r save clustering fm, cache = F}
if (params$fm == "y"){
  saveRDS(clustering, "clustering_flowmeans.rds")
}
```

```{r flowmeans matching}
if (params$fm == "y"){
  # MATCHING LABELS AND PREDICTIONS
  matched <- matching(clustering, labels, threshold)
}
```

```{r save matching fm, cache = F}
if (params$fm == "y"){
  saveRDS(matched, "matched_flowmeans.rds")
}
```

```{r flowmeans performance}
if (params$fm == "y"){
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                        reference = labels[labels != "outliers"])
  cm_merged <- confusionMatrix(
    data = matched$merged_clusters[matched$merged_labels != "outliers"],
    reference = matched$merged_labels[matched$merged_labels != "outliers"]
  )
  
  # COMPUTE MEAN F1 and WEIGHTED MEAN F1
  mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
}
```

```{r flowmeans results, cache = FALSE}
if (params$fm == "y"){
  summ[row, c(1,2,4)] <- c(params$id, params$norm, method)
  summ[row, -c(1,2,4)] <- c(nrow(ff@exprs), threshold, length(table(clustering)),
                         matched$partitions,
                         mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                         NA, NA, ptm[1] + ptm[4], ptm[3])
  row <- row + 1
  
  # FOR VISUALIZATION
  cat("\n", method)
  cat("\nContingency matrix")
  print(matched$c)
  cat("\nF1 matrix")
  print(round(matched$f, 2))
  cat("\nConfusion matrix")
  print(cm$table)
  cat("\nConfusion matrix (MERGED)")
  print(cm_merged$table)
}
```

##  flowPeaks
```{r flowpeaks clustering}
if (params$fp == "y"){
  method <- "flowPeaks"
  
  # clustering
  ptm <- proc.time() # measuring CPU time
  pred <- flowPeaks(ff@exprs)
  ptm <- proc.time() - ptm
  
  clustering <- pred$peaks.cluster
}
```

```{r save clustering fp, cache = F}
if (params$fp == "y"){
  saveRDS(clustering, "clustering_flowpeaks.rds")
}
```

```{r flowpeaks matching}
if (params$fp == "y"){
  # MATCHING LABELS AND PREDICTIONS
  matched <- matching(clustering, labels, threshold)
}
```

```{r save matching fp, cache = F}
if (params$fp == "y"){
  saveRDS(matched, "matched_flowpeaks.rds")
}
```

```{r flowpeaks performance}
if (params$fp == "y"){
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                        reference = labels[labels != "outliers"])
  cm_merged <- confusionMatrix(
    data = matched$merged_clusters[matched$merged_labels != "outliers"],
    reference = matched$merged_labels[matched$merged_labels != "outliers"]
  )
  
  # COMPUTE MEAN F1 and WEIGHTED MEAN F1
  mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
}
```

```{r flowpeaks results, cache = FALSE}
if (params$fp == "y"){
  summ[row, c(1,2,4)] <- c(params$id, params$norm, method)
  summ[row, -c(1,2,4)] <- c(nrow(ff@exprs), threshold, length(table(clustering)),
                          matched$partitions,
                          mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                          NA, NA, ptm[1] + ptm[4], ptm[3])
  row <- row + 1
  
  # FOR VISUALIZATION
  cat("\n", method)
  cat("\nContingency matrix")
  print(matched$c)
  cat("\nF1 matrix")
  print(round(matched$f, 2))
  cat("\nConfusion matrix")
  print(cm$table)
  cat("\nConfusion matrix (MERGED)")
  print(cm_merged$table)
}
```

##  DEPECHE
```{r depeche clustering}
if (params$dp == "y"){
  method <- "DEPECHE"
  
  set.seed(42)
  # clustering
  ptm <- proc.time() # measuring CPU time
  pred <-  depeche(ff@exprs, createOutput = F)
  ptm <- proc.time() - ptm
  
  clustering <- pred$clusterVector
}
```

```{r save clustering dp, cache = F}
if (params$dp == "y"){
  saveRDS(clustering, "clustering_depeche.rds")
}
```

```{r depeche matching, eval = F}
if (params$dp == "y"){
  # MATCHING LABELS AND PREDICTIONS
  matched <- matching(clustering, labels, threshold)
}
```

```{r save matching dp, cache = F}
if (params$dp == "y"){
  saveRDS(matched, "matched_depeche.rds")
}
```

```{r depeche performance}
if (params$dp == "y"){
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                        reference = labels[labels != "outliers"])
  cm_merged <- confusionMatrix(
    data = matched$merged_clusters[matched$merged_labels != "outliers"],
    reference = matched$merged_labels[matched$merged_labels != "outliers"]
  )
  
  # COMPUTE MEAN F1 and WEIGHTED MEAN F1
  mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
}
```

```{r depeche results, cache = FALSE}
if (params$dp == "y"){
  summ[row, c(1,2,4)] <- c(params$id, params$norm, method)
  summ[row, -c(1,2,4)] <- c(nrow(ff@exprs), threshold, length(table(clustering)),
                          matched$partitions,
                          mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                          NA, NA, ptm[1] + ptm[4], ptm[3])
  row <- row + 1
  
  # FOR VISUALIZATION
  cat("\n", method)
  cat("\nContingency matrix")
  print(matched$c)
  cat("\nF1 matrix")
  print(round(matched$f, 2))
  cat("\nConfusion matrix")
  print(cm$table)
  cat("\nConfusion matrix (MERGED)")
  print(cm_merged$table)
}
```

# Clustering on t-SNE-reduced data (2D)
##  RphenoGraph
```{r tsne phenograph clustering}
if (params$tpg == "y"){
  method <- "t-SNE + PhenoGraph"
  
  # clustering
  ptm <- proc.time() # measuring CPU time
  pred <- Rphenograph(tsne_red) # default k = 30 nearest neighbours
  ptm <- proc.time() - ptm
  
  clustering <- pred$membership
}
```

```{r save clustering tsne pheno, cache = F}
if (params$tpg == "y"){
  saveRDS(clustering, "clustering_t_pheno.rds")
}
```

```{r tsne phenograph matching}
if (params$tpg == "y"){
  # MATCHING LABELS AND PREDICTIONS
  matched <- matching(clustering, labels, threshold)
}
```

```{r save matching tsne pheno, cache = F}
if (params$tpg == "y"){
  saveRDS(matched, "matched_t_pheno.rds")
}
```

```{r tsne phenograph performance}
if (params$tpg == "y"){
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                        reference = labels[labels != "outliers"])
  cm_merged <- confusionMatrix(
    data = matched$merged_clusters[matched$merged_labels != "outliers"],
    reference = matched$merged_labels[matched$merged_labels != "outliers"]
  )
  
  # COMPUTE MEAN F1 and WEIGHTED MEAN F1
  mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
}
```

```{r tsne phenograph results, cache = FALSE}
if (params$tpg == "y"){
  summ[row, c(1,2,4)] <- c(params$id, params$norm, method)
  summ[row, -c(1,2,4)] <- c(nrow(ff@exprs), threshold, length(table(clustering)),
                          matched$partitions,
                          mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                          tsne_time[1] + tsne_time[4], tsne_time[3], ptm[1] + ptm[4], ptm[3])
  row <- row + 1
  
  # FOR VISUALIZATION
  cat("\n", method)
  cat("\nContingency matrix")
  print(matched$c)
  cat("\nF1 matrix")
  print(round(matched$f, 2))
  cat("\nConfusion matrix")
  print(cm$table)
  cat("\nConfusion matrix (MERGED)")
  print(cm_merged$table)
}
```

## flowMeans
```{r tsne flowmeans clustering}
if (params$tfm == "y"){
  method <- "t-SNE + flowMeans"
  
  # clustering
  ptm <- proc.time() # measuring CPU time
  pred <- flowMeans(tsne_red, MaxN = n_pops )
  ptm <- proc.time() - ptm
  
  clustering <- pred@Labels[[1]]
}
```

```{r save clustering tsne fm, cache = F}
if (params$tfm == "y"){
  saveRDS(clustering, "clustering_t_flowmeans.rds")
}
```

```{r tsne flowmeans matching}
if (params$tfm == "y"){
  # MATCHING LABELS AND PREDICTIONS
  matched <- matching(clustering, labels, threshold)
}
```

```{r save matching tsne fm, cache = F}
if (params$tfm == "y"){
  saveRDS(matched, "matched_t_flowmeans.rds")
}
```

```{r tsne flowmeans performance}
if (params$tfm == "y"){
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                        reference = labels[labels != "outliers"])
  cm_merged <- confusionMatrix(
    data = matched$merged_clusters[matched$merged_labels != "outliers"],
    reference = matched$merged_labels[matched$merged_labels != "outliers"]
  )
  
  # COMPUTE MEAN F1 and WEIGHTED MEAN F1
  mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
}
```

```{r tsne flowmeans results, cache = FALSE}
if (params$tfm == "y"){
  summ[row, c(1,2,4)] <- c(params$id, params$norm, method)
  summ[row, -c(1,2,4)] <- c(nrow(ff@exprs), threshold, length(table(clustering)),
                          matched$partitions,
                          mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                          tsne_time[1] + tsne_time[4], tsne_time[3], ptm[1] + ptm[4], ptm[3])
  row <- row + 1
  
  # FOR VISUALIZATION
  cat("\n", method)
  cat("\nContingency matrix")
  print(matched$c)
  cat("\nF1 matrix")
  print(round(matched$f, 2))
  cat("\nConfusion matrix")
  print(cm$table)
  cat("\nConfusion matrix (MERGED)")
  print(cm_merged$table)
}
```

## flowPeaks
```{r tsne flowpeaks clustering}
if (params$tfp == "y"){
  method <- "t-SNE + flowPeaks"
  
  # clustering
  ptm <- proc.time() # measuring CPU time
  pred <- flowPeaks(tsne_red) 
  ptm <- proc.time() - ptm
  
  clustering <- pred$peaks.cluster
}
```

```{r save clustering tsne fp, cache = F}
if (params$tfp == "y"){
  saveRDS(clustering, "clustering_t_flowpeaks.rds")
}
```

```{r tsne flowpeaks matching}
if (params$tfp == "y"){
  # MATCHING LABELS AND PREDICTIONS
  matched <- matching(clustering, labels, threshold)
}
```

```{r save matching tsne fp}
if (params$tfp == "y"){
  saveRDS(matched, "matched_t_flowpeaks.rds")
}
```

```{r tsne flowpeaks performance}
if (params$tfp == "y"){
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                        reference = labels[labels != "outliers"])
  cm_merged <- confusionMatrix(
    data = matched$merged_clusters[matched$merged_labels != "outliers"],
    reference = matched$merged_labels[matched$merged_labels != "outliers"]
  )
  
  # COMPUTE MEAN F1 and WEIGHTED MEAN F1
  mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
}
```

```{r tsne flowpeaks results, cache = FALSE}
if (params$tfp == "y"){
  summ[row, c(1,2,4)] <- c(params$id, params$norm, method)
  summ[row, -c(1,2,4)] <- c(nrow(ff@exprs), threshold, length(table(clustering)),
                          matched$partitions,
                          mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                          tsne_time[1] + tsne_time[4], tsne_time[3], ptm[1] + ptm[4], ptm[3])
  row <- row + 1
  
  # FOR VISUALIZATION
  cat("\n", method)
  cat("\nContingency matrix")
  print(matched$c)
  cat("\nF1 matrix")
  print(round(matched$f, 2))
  cat("\nConfusion matrix")
  print(cm$table)
  cat("\nConfusion matrix (MERGED)")
  print(cm_merged$table)
}
```

##  DEPECHE
```{r tsne depeche clustering}
if (params$tdp == "y"){
  method <- "t-SNE + DEPECHE"
  
  set.seed(42)
  # clustering
  ptm <- proc.time() # measuring CPU time
  pred <-  depeche(tsne_red, createOutput = F)
  ptm <- proc.time() - ptm
  
  clustering <- pred$clusterVector
}
```

```{r save clustering tsne depeche}
if (params$tdp == "y"){
  saveRDS(clustering, "clustering_t_depeche.rds")
}
```

```{r tsne depeche matching}
if (params$tdp == "y"){
  # MATCHING LABELS AND PREDICTIONS
  matched <- matching(clustering, labels, threshold)
}
```

```{r save matching tsne depeche}
if (params$tdp == "y"){
  saveRDS(matched, "matched_t_depeche.rds")
}
```

```{r tsne depeche performance}
if (params$tdp == "y"){
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                        reference = labels[labels != "outliers"])
  cm_merged <- confusionMatrix(
    data = matched$merged_clusters[matched$merged_labels != "outliers"],
    reference = matched$merged_labels[matched$merged_labels != "outliers"]
  )
  
  # COMPUTE MEAN F1 and WEIGHTED MEAN F1
  mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
}
```

```{r tsne depeche results, cache = FALSE}
if (params$tdp == "y"){
  summ[row, c(1,2,4)] <- c(params$id, params$norm, method)
  summ[row, -c(1,2,4)] <- c(nrow(ff@exprs), threshold, length(table(clustering)),
                          matched$partitions,
                          mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                          tsne_time[1] + tsne_time[4], tsne_time[3], ptm[1] + ptm[4], ptm[3])
  row <- row + 1
  
  # FOR VISUALIZATION
  cat("\n", method)
  cat("\nContingency matrix")
  print(matched$c)
  cat("\nF1 matrix")
  print(round(matched$f, 2))
  cat("\nConfusion matrix")
  print(cm$table)
  cat("\nConfusion matrix (MERGED)")
  print(cm_merged$table)
}
```

##  ClusterX
```{r tsne clusterx clustering}
if (params$tcx == "y"){
  method <- "t-SNE + ClusterX"
  
  # clustering
  ptm <- proc.time() # measuring CPU time
  clustering <- cytof_cluster(tsne_red, method = "ClusterX")
  ptm <- proc.time() - ptm
}
```

```{r save clustering tsne clusterx}
if (params$tcx == "y"){
  saveRDS(clustering, "clustering_t_clusterx.rds")
}
```

```{r tsne clusterx matching}
if (params$tcx == "y"){
  # MATCHING LABELS AND PREDICTIONS
  matched <- matching(clustering, labels, threshold)
}
```

```{r save matching tsne clusterx}
if (params$tcx == "y"){
  saveRDS(matched, "matched_t_clusterx.rds")
}
```

```{r tsne clusterx performance}
if (params$tcx == "y"){
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                        reference = labels[labels != "outliers"])
  cm_merged <- confusionMatrix(
    data = matched$merged_clusters[matched$merged_labels != "outliers"],
    reference = matched$merged_labels[matched$merged_labels != "outliers"]
  )
  
  # COMPUTE MEAN F1 and WEIGHTED MEAN F1
  mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
}
```

```{r tsne clusterx results, cache = FALSE}
if (params$tcx == "y"){
  summ[row, c(1,2,4)] <- c(params$id, params$norm, method)
  summ[row, -c(1,2,4)] <- c(nrow(ff@exprs), threshold, length(table(clustering)),
                          matched$partitions,
                          mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                          tsne_time[1] + tsne_time[4], tsne_time[3], ptm[1] + ptm[4], ptm[3])
  row <- row + 1
  
  # FOR VISUALIZATION
  cat("\n", method)
  cat("\nContingency matrix")
  print(matched$c)
  cat("\nF1 matrix")
  print(round(matched$f, 2))
  cat("\nConfusion matrix")
  print(cm$table)
  cat("\nConfusion matrix (MERGED)")
  print(cm_merged$table)
}
```

# Clustering on UMAP-reduced data (2D)
##  RphenoGraph
```{r umap phenograph clustering}
if (params$upg == "y"){
  method <- "UMAP + PhenoGraph"
  
  # clustering
  ptm <- proc.time() # measuring CPU time
  pred <- Rphenograph(umap_red$layout) # default k = 30 nearest neighbours
  ptm <- proc.time() - ptm
  
  clustering <- pred$membership
}
```

```{r save clustering umap pheno}
if (params$upg == "y"){
  saveRDS(clustering, "clustering_u_pheno.rds")
}
```

```{r umap phenograph matching}
if (params$upg == "y"){
  # MATCHING LABELS AND PREDICTIONS
  matched <- matching(clustering, labels, threshold)
}
```

```{r save matching umap pheno}
if (params$upg == "y"){
  saveRDS(matched, "matched_u_pheno.rds")
}
```

```{r umap phenograph performance}
if (params$upg == "y"){
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                        reference = labels[labels != "outliers"])
  cm_merged <- confusionMatrix(
    data = matched$merged_clusters[matched$merged_labels != "outliers"],
    reference = matched$merged_labels[matched$merged_labels != "outliers"]
  )
  
  # COMPUTE MEAN F1 and WEIGHTED MEAN F1
  mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
}
```

```{r umap phenograph results, cache = FALSE}
if (params$upg == "y"){
  summ[row, c(1,2,4)] <- c(params$id, params$norm, method)
  summ[row, -c(1,2,4)] <- c(nrow(ff@exprs), threshold, length(table(clustering)),
                          matched$partitions,
                          mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                          umap_time[1] + umap_time[4], umap_time[3], ptm[1] + ptm[4], ptm[3])
  row <- row + 1
  
  # FOR VISUALIZATION
  cat("\n", method)
  cat("\nContingency matrix")
  print(matched$c)
  cat("\nF1 matrix")
  print(round(matched$f, 2))
  cat("\nConfusion matrix")
  print(cm$table)
  cat("\nConfusion matrix (MERGED)")
  print(cm_merged$table)
}
```

##  flowMeans
```{r umap flowmeans clustering}
if (params$ufm == "y"){
  method <- "UMAP + flowMeans"
  
  # clustering
  ptm <- proc.time() # measuring CPU time
  pred <- flowMeans(umap_red$layout, MaxN = n_pops )
  ptm <- proc.time() - ptm
  
  clustering <- pred@Labels[[1]]
}
```

```{r save clustering umap fm}
if (params$ufm == "y"){
  saveRDS(clustering, "clustering_u_flowmeans.rds")
}
```

```{r umap flowmeans matching}
if (params$ufm == "y"){
# MATCHING LABELS AND PREDICTIONS
  matched <- matching(clustering, labels, threshold)
}
```

```{r save matching umap flowmean}
if (params$ufm == "y"){
  saveRDS(matched, "matched_u_flowmeans.rds")
}  
```

```{r umap flowmeans performance}
if (params$ufm == "y"){
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                        reference = labels[labels != "outliers"])
  cm_merged <- confusionMatrix(
    data = matched$merged_clusters[matched$merged_labels != "outliers"],
    reference = matched$merged_labels[matched$merged_labels != "outliers"]
  )
  
  # COMPUTE MEAN F1 and WEIGHTED MEAN F1
  mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
}
```

```{r umap flowmeans results, cache = FALSE}
if (params$ufm == "y"){
  summ[row, c(1,2,4)] <- c(params$id, params$norm, method)
  summ[row, -c(1,2,4)] <- c(nrow(ff@exprs), threshold, length(table(clustering)),
                          matched$partitions,
                          mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                          umap_time[1] + umap_time[4], umap_time[3], ptm[1] + ptm[4], ptm[3])
  row <- row + 1
  
  # FOR VISUALIZATION
  cat("\n", method)
  cat("\nContingency matrix")
  print(matched$c)
  cat("\nF1 matrix")
  print(round(matched$f, 2))
  cat("\nConfusion matrix")
  print(cm$table)
  cat("\nConfusion matrix (MERGED)")
  print(cm_merged$table)
}
```

##  flowPeaks
```{r umap flowpeaks clustering}
if (params$ufp == "y"){
  method <- "UMAP + flowPeaks"
  
  # clustering
  ptm <- proc.time() # measuring CPU time
  pred <- flowPeaks(umap_red$layout)
  ptm <- proc.time() - ptm
  
  clustering <- pred$peaks.cluster
}
```

```{r save clustering umap fp}
if (params$ufp == "y"){
  saveRDS(clustering, "clustering_u_flowpeaks.rds")
}
```

```{r umap flowpeaks matching}
if (params$ufp == "y"){
  # MATCHING LABELS AND PREDICTIONS
  matched <- matching(clustering, labels, threshold)
}
```

```{r save matching umap flowpeaks}
if (params$ufp == "y"){
  saveRDS(matched, "matched_u_flowpeaks.rds")
}
```

```{r umap flowpeaks performance}
if (params$ufp == "y"){
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                        reference = labels[labels != "outliers"])
  cm_merged <- confusionMatrix(
    data = matched$merged_clusters[matched$merged_labels != "outliers"],
    reference = matched$merged_labels[matched$merged_labels != "outliers"]
  )
  
  # COMPUTE MEAN F1 and WEIGHTED MEAN F1
  mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
}
```

```{r umap flowpeaks results, cache = FALSE}
if (params$ufp == "y"){
  summ[row, c(1,2,4)] <- c(params$id, params$norm, method)
  summ[row, -c(1,2,4)] <- c(nrow(ff@exprs), threshold, length(table(clustering)),
                          matched$partitions,
                          mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                          umap_time[1] + umap_time[4], umap_time[3], ptm[1] + ptm[4], ptm[3])
  row <- row + 1
  
  # FOR VISUALIZATION
  cat("\n", method)
  cat("\nContingency matrix")
  print(matched$c)
  cat("\nF1 matrix")
  print(round(matched$f, 2))
  cat("\nConfusion matrix")
  print(cm$table)
  cat("\nConfusion matrix (MERGED)")
  print(cm_merged$table)
}
```

##  DEPECHE
```{r umap depeche clustering}
if (params$udp == "y"){
  method <- "UMAP + DEPECHE"
  
  set.seed(42)
  # clustering
  ptm <- proc.time() # measuring CPU time
  pred <-  depeche(umap_red$layout, createOutput = F)
  ptm <- proc.time() - ptm
  
  clustering <- pred$clusterVector
}
```

```{r save clustering umap depeche}
if (params$udp == "y"){
  saveRDS(clustering, "clustering_u_depeche.rds")
}
```

```{r umap depeche matching}
if (params$udp == "y"){
  # MATCHING LABELS AND PREDICTIONS
  matched <- matching(clustering, labels, threshold)
}
```

```{r save matching umap depeche}
if (params$udp == "y"){
  saveRDS(matched, "matched_u_depeche.rds")
}
```

```{r umap depeche performance}
if (params$udp == "y"){
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                        reference = labels[labels != "outliers"])
  cm_merged <- confusionMatrix(
    data = matched$merged_clusters[matched$merged_labels != "outliers"],
    reference = matched$merged_labels[matched$merged_labels != "outliers"]
  )
  
  # COMPUTE MEAN F1 and WEIGHTED MEAN F1
  mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
}
```

```{r umap depeche results, cache = FALSE}
if (params$udp == "y"){
  summ[row, c(1,2,4)] <- c(params$id, params$norm, method)
  summ[row, -c(1,2,4)]<- c(nrow(ff@exprs), threshold, length(table(clustering)),
                         matched$partitions,
                         mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                         umap_time[1] + umap_time[4], umap_time[3], ptm[1] + ptm[4], ptm[3])
  row <- row + 1
  
  # FOR VISUALIZATION
  cat("\n", method)
  cat("\nContingency matrix")
  print(matched$c)
  cat("\nF1 matrix")
  print(round(matched$f, 2))
  cat("\nConfusion matrix")
  print(cm$table)
  cat("\nConfusion matrix (MERGED)")
  print(cm_merged$table)
}
```

##  ClusterX
```{r umap clusterx clustering}
if (params$ucx == "y"){
  method <- "UMAP + ClusterX"
  
  # clustering
  ptm <- proc.time() # measuring CPU time
  clustering <- cytof_cluster(umap_red$layout, method = "ClusterX")
  ptm <- proc.time() - ptm
}
```

```{r save clustering umap clusterx}
if (params$ucx == "y"){
saveRDS(clustering, "clustering_u_clusterx.rds")
}
```

```{r umap clusterx matching}
if (params$ucx == "y"){
  # MATCHING LABELS AND PREDICTIONS
  matched <- matching(clustering, labels, threshold)
}
```

```{r save matching umap clusterx}
if (params$ucx == "y"){
  saveRDS(matched, "matched_u_clusterx.rds")
}
```

```{r umap clusterx performance}
if (params$ucx == "y"){
  # COMPUTING THE CONFUSION MATRIX AND OTHER PERFORMANCE MEASUREMENTS
  cm <- confusionMatrix(data = matched$clusters[labels != "outliers"],
                        reference = labels[labels != "outliers"])
  cm_merged <- confusionMatrix(
    data = matched$merged_clusters[matched$merged_labels != "outliers"],
    reference = matched$merged_labels[matched$merged_labels != "outliers"]
  )
  
  # COMPUTE MEAN F1 and WEIGHTED MEAN F1
  mf1 <- mean_f1(cm, cm_merged, matched$merged_labels, matched$count_merged_pops)
}
```

```{r umap clusterx results, cache = FALSE}
if (params$ucx == "y"){
  summ[row, c(1,2,4)] <- c(params$id, params$norm, method)
  summ[row, -c(1,2,4)]<- c(nrow(ff@exprs), threshold, length(table(clustering)),
                         matched$partitions,
                         mf1$mf1, mf1$mf1_w, mf1$mf1_i_w, mf1$mf1_m, mf1$mf1_m_c, 
                         umap_time[1] + umap_time[4], umap_time[3], ptm[1] + ptm[4], ptm[3])
  row <- row + 1
  
  # FOR VISUALIZATION
  cat("\n", method)
  cat("\nContingency matrix")
  print(matched$c)
  cat("\nF1 matrix")
  print(round(matched$f, 2))
  cat("\nConfusion matrix")
  print(cm$table)
  cat("\nConfusion matrix (MERGED)")
  print(cm_merged$table)
}
```

# Export results
```{r export results, cache = FALSE}
# export results 
write.table(summ, "summary.txt", append = T, sep = "\t", row.names = F, col.names = T)
```

# Libraries
```{r sessioninfo, cache = F}
sessionInfo()
```













